### 2018-10-04 slides/ERE_lunch.Rmd
Finally got the redistribution data ready. Will run spatial analyses tomorrow

---
title: "Displacement of fishing effort by Large Scale Marine Protected Areas"
author:
  - Juan Carlos Villaseñor-Derbez^1^
  - John Lynham^2^
institute:
  - ^1^Bren School of Environmental Science and Management, UC Santa Barbara
  - ^2^Department of Economics, University of Hawaii at Manoa
date: "5/10/2018"
output: 
  beamer_presentation: 
    toc: yes
    slide_level: 2
    citation_package: biblatex
bibliography: references.bib
biblio-style: authoryear
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      eval = TRUE)

suppressPackageStartupMessages({
  library(startR)
  library(sf)
  library(tidyverse)
})

source(here::here("scripts", "st_rotate.R"))
```

# Background

## MPAs

- Marine Protected Areas (MPAs): Spatial management of fishing effort
-Areas > 250,000 Km^2^ are Large Scale (LSMPAs) \footcite{toonen_2013}
  - Implemented in the pelagic environment
  - Dominant human activity is industrial fishing \footcite{gray_2017}

```{r}
mpas <- sf::read_sf(here::here("raw_data", "spatial", "WDPA_Mar2018"),
        layer = "WDPA_Mar2018_marine-shapefile-polygons",
        quiet = T,
        stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  filter(gis_m_area > 250000)

coast <- rnaturalearth::ne_countries(returnclass = "sf")
```

```{r, fig.height = 3.5}
ggplot() +
  geom_sf(data = coast) +
  geom_sf(data = mpas, fill = "red") +
  ggtheme_plot()
```

## PIPA

- Phoenix Island Protected Area
  - Part of Kiribati EEZ
  - Implemented in 2015
  - Kiribati is part of the PNA, along with other 8 countries

```{r}
eez <- sf::read_sf(dsn = here::here("data", "spatial", "PNA_EEZ"), layer = "PNA_EEZ") %>% 
  sf::st_transform(crs = "+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_rotate() %>% 
  st_buffer(0) %>% 
  group_by(ISO_Ter1) %>% 
  summarize()

pipa <- sf::read_sf(dsn = here::here("data", "spatial", "PIPA"), layer = "PIPA") %>% 
  sf::st_transform(crs = "+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_rotate()
```

```{r, fig.height = 4}
ggplot() +
  geom_sf(data = eez, fill = "gray", alpha = 0.5) +
  geom_sf(data = pipa, fill = "red") +
  ggtheme_plot()
```

## PIPA

- Phoenix Island Protected Area
  - Part of Kiribati EEZ
  - Implemented in 2015
  - Kiribati is part of the PNA, along with other 8 countries

- Recent focus of *Blue paradox* \footcite{mcdermott_2018}

```{r, fig.height = 4}
displacement_data <- readRDS(file = here::here("data", "displacement_data.rds")) %>% 
  filter(year < 2017)

displacement_data %>% 
  filter(gear == "purse_seines",
         country == "PIPA") %>% 
  ggplot(aes(x = date, y = h_prop, group = post)) +
  geom_smooth(method = "loess", color = "black", linetype = "dashed") +
  geom_point(aes(color = post), size = 2) +
  geom_vline(xintercept = lubridate::date(paste(2015, 1, 1, sep = "/")), linetype = "dashed") +
  scale_fill_brewer(palette = "Paired") +
  ggtheme_plot() +
  ylab("Proportion of total hours by vessels") +
  scale_y_continuous(labels = scales::percent) +
  scale_color_brewer(palette = "Set1")
```

```{r}
effort_by_vessel <- readRDS(file = here::here("data", "effort_by_vessel.rds")) %>% 
  filter(year < 2018)
```

# Objectives

- How do vessel-level behavior change due to PIPA implementation?
- What happens to the displaced fishing effort?

# Methods

## Data

## Analyses

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \mu_1Y_t + \mu_2Y_t^2 + \phi_t + \gamma_i + \epsilon_{i,t}
$$

$$
y_{i,t} = \alpha + \beta_1Post_t + \beta_{2,i}Country_i + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
$$

# Preliminary results

# Future work

# References