### 2018-10-17 docs/Manuscript.tex
Updated manuscript intro and methods

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Displacement of fishing effort by Large Scale Marine Protected Areas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{chicago}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Displacement of fishing effort by Large Scale Marine Protected Areas}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-16}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

\usepackage{float}
\floatplacement{figure}{H}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\begin{document}
\maketitle
\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, and include increases in
abundance, biomass, and diversity within the bounded regions. These
no-take zones may produce spillover effects, which provide fish for
outside areas. However, the economic and ecological implications of
displacing fishing effort are not yet fully understood. Novel data
products that track fishing effort at the vessel-level allow us to
identify changes in fleet- and vessel-level behavior upon the
implementation of protected areas, as well as how these redistribute.
This papers evaluates the implications of implementing LSMPA, by
evaluating changes in fishing hours, showing that vessels in the
effected region reduce fishing effort after the implementation of PIPA.
Our results are robust to a set of specifications. We also track the
relative spatial allocation of fishing events thorugh time, and identify
that areas closer to PIPA show an increase in relative fishing hourse
due to the displacement of PIPA-fishing vessels. Our results not only
provide an impact evaluation of the effect of LSMPAs on fishing
activity, but provide insights into vessel redistribution dynamics,
which may have ecological and economic implications.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environments by
2020. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018}. A significant part
of this rapid increase can be attributed to the designation of a small
number of Large Scale Marine Protected Areas (LSMPAs;
\citet{boonzaier_2016,singleton_2014,gray_2017}). These are defined as
MPAS larger than 250,000 Km\textsuperscript{2} in extension (Fig.
\ref{fig:LSMPAs_map}), and are often implemented in the pelagic
environment, where the dominant human activity is industrial fishing
\citep{toonen_2013,gray_2017,kroodsma_2018}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. Moreover, the world has seen a widespread increase in
the implementation of LSMPAs, likely as a way to rapidly achieve the
10\% target \citep{boonzaier_2016,alger_2017,sala_2018}. However, as
with customary MPAs, it is important that we understand the
socioeconomic implications of management interventions that result in
the spatial displacement of fishing effort. They \textbf{key gap in the
literature is X}. We adress it here by using new stuff and find results
that are of relevance to policy makers and advance our knowledge, which
should be incorporated when designing and planning for LSMPAs.

Due to weak property rights, limited habitat transformation, and
potentially lower management costs, pelagic MPAs provide an opportunity
to safeguard the oceans \citet{game_2009}. A growing body of literature
has shown that closing the high seas to all fishing could increase
fishery yields and profitability of fisheries, with negligible costs to
food security \citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}.

The early literature on LSMPAs focused on the inherent difficulties that
come with a pelagic (\emph{i.e.} open water) environment, where
organisms typically have large home ranges. \citet{game_2009} suggest
that most of the challenges can be overcome with the incorporation of
technology, in what then became known as Dynamic Ocean Management
\citep{maxwell_2015}. \citet{kaplan_2009} claim that very large MPAs
would result in excessive opportunity costs and that these would be
difficult to enforce. \citet{toonen_2013} show that just a small number
of LSMPAs made up 80\% of the managed areas in the ocean in 2013, and
that these can help achieve the CBD Aichi targets faster.
\citet{singleton_2014} provide an objective discussion of pros and cons
of LSMPAs.

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these human
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protectes two
thrids of the tagged grey reef sharks by effectively excluding fishing
effort. No studies have evaluated the displacement of fishing effort due
to LSMPA implementation.

The exclusion of these LSMPAs is likely to change fisher's behavior.
Theoretical models of fishing effort redistribution range from the
simplistic assumption that effort inside the bounded region disappears,
to spatially explicit models that reallocate fishing effort based on
habitat characteristics, presence of other vessels, and expected returns
\citep{smith_2003,hilborn_2006}. However, these focus on the long term
optimal equilibrium and redistribution of fishing effort may not be
optimal, especially over the first years \citep{stevenson_2013}. The few
empirical works suggest that resource users may show idiosyncratic
responses. For example, \citet{stevenson_2013} show that a network of
MPAs displaced fishing effort farther away from ports resulting in
higher \emph{perceived} costs, but that catch per unit effort also
increased. \citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of MPAs in California,
and find that commercial dive boats follow a fishing-the-line pattern,
while some fishing boats follow an ideal free distribution. The way in
which fishers react to a spatial closure can have major implications in
its outcome \citep{smith_2003,hilborn_2006} highlighting the need to
understand how fishers react to the implementation of LSMPAs, and
fishing effort changes and is spatially redistributed.

The main objective of this paper is to identify behavioral responses to
the implementation of LSMPAs. We combine novel vessel tracking
technologies and causal inference techniques to identify behavioral
changes of fishing vessels due to the implementation of PIPA. We focus
on fishing hours and distance traveled as outcome variables that fishers
might adjust following implementation of a LSMPA in an impact-evaluation
fashion. Additionally, we evaluate the spatial redistribution of fishing
effort that existed within PIPA before its implementation. This work
provides novel empirical insights into fisher's responses to the
implementation of LSMPAs, and can help guide future interventions. Our
work is novel in the sense that it provides empirical evidence of the
effect of large-acale Marine Protected Areas in fishing behavior and
distribution.

The next sections are as follows: Section \ref{background} provides an
overview of the Nauru Agreement and associated countries, a description
of the fleet that operates in the region, and a brief history of PIPA.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, section \ref{discussion}
provides an extension of our results to other LSMPAs and discusses our
results.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated english name is `Protected Area'.}
\end{figure}

\hypertarget{background}{%
\section{Background}\label{background}}

The Nauru Agreement was established in 1982 by Pacific island nations to
manage their important tuna resources. PNA Members include Federated
States of Micronesia, Kiribati, Marshall Islands, Nauru, Palau, Papua
New Guinea, Solomon Islands, and Tuvalu.The Nauru Agreement regulated
access of foreign vessels (\emph{i.e.} those from non-PNA countries).
Holding \textasciitilde{}80\% of the historical purse seining grounds
within their Exclusive Economic Zones, PNA countries gained bargaining
power when providing access to foreign fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisors recomend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. A solid line indicates
countries that belong to the PNA, while a dashed line indicates all
others. A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. The pacific region
has historically accounted for a large portion of tuna species
production \citep{aqorau_1997}. Today, the PNA controls close to 50\% of
the global skipjack tuna production \citep{pna_website_2018}. A large
portion of these catches derive from purse seine vessels licensed under
the VDS. Fishing vessels from Australia, New Zeland, China, France,
Korea, Japan, the Philippines, Taiwan, and the United States participate
in the purse-seining VDS.

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Island Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of it was declared as no-take. In
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 Km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. The implementation of such a large
conservation area in one of the most important fishing regions in the
world provides a great opportunity to evaluate the behavioral responses
and redistribution of fishing effort by vessels that used to fish there.
\citet{mcdermott_2018} showed that fishing effort within the Phoenix
Islands Marine Protected Area (PIPA) increased between the anouncement
(September 1\textsuperscript{st}, 2014) and its implementation in
Janyary 1\textsuperscript{st}, 2015. Likewise, they demonstrate that
fishing effort is effectively reduced after implementation. To this, we
pose two questions: How do individual vessels respond to the sudden
exclusion of such a big area? And where did all the vessels go? In the
next sections we describe the data and methods used to answer this
questions.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

This section is divided into two main parts. First, we provide a general
description of AIS data and the process of identification of
vessel-level fishing events done by Global Fishing Watch\footnote{Global
  Fishing Watch: \url{globalfishingwatch.org}}. Alongside, we describe
the subset of data that we use for these analyses. When relevant, we
also point out possible shortcomings in the data, or factors that must
be considered in the later analyses. We then move on to explain our
empirical strategy for the identification of the behavioral changes and
redistribution of fishing effort.

\hypertarget{data}{%
\subsection{Data}\label{data}}

Automatic Identification Systems are on-board devices intended to
provide at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activities to surrounding vessels. These
broadcasted messages can be received by satellites and land-based
antennas. GFW uses a neural network to infer vessel characteristics and
whether each broadcasted position represents a fishing event, thus
allowing us to estimate near real-time fishing events globally since
2012 \citep{kroodsma_2018}. Our data contain information for 2012 -
2017. The recent addition of satellites that can receive AIS signals
causes an apparent increase in the number of broadcasted AIS messages
(\emph{i.e.} points), and therefore number of vessels and fishing hours.
The variability in AIS data and ocean conditions require that temporal
trends be taken into account. We do that by obtaining a subset of data
that meet a BACI design, which gives us the full tracks for vessels
affected and unaffected by the implementation of PIPA.

Our data contain over 45 million individual AIS messages (\emph{i.e.}
positions) for 371 fishing vessels that at some point fished in PNA
waters. A total of 233 vessels have fished within PIPA waters; 217 did
so at least once before 2015. However, not all vessels continued to fish
elsewhere after PIPA implementation: 34 vessels have no recorded AIS
messages after 2015, leaving us with 176 vessels that fished inside PIPA
before its implementation, and continued to fish elsewhere
afterward\footnote{The 34 missing vessels might have exited the fishery, been decommissioned or sold (therefore changing their AIS and mmsi), or turned off their AIS transmitters. In either case, we are not able to observe these.}.
From these, we focus on purse seiners -the most important for PNA
countries- and are left with 61
vessels\footnote{We perform some of the same analyses for longliners and include them in the Appendix}.
New vessels might have also entered the fishery after PIPA closure, and
were likely not exposed to the policy intervention in the pre-treatment
period. To account for this, we identify a subset of vessels which we
track since before the implementation of PIPA, and categorize them as
treated or control vessels. Our treatment and control groups are defined
as follows.

The treatment group contains all purse seiners (n = 61) that fished
within PIPA at least once before the anouncement (09/01/2014), and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure. This definition of control vessels provides us with 26 purse
seiners that never fished inside PIPA but fished in other PNA waters
before and after PIPA implementation. During our causal identification,
we also use two more strict definitions of control groups: one with only
vessels that belong to PNA countries (n = 7), and one that excludes
Chinese vessels (n = 21). Our definition of treatment and control groups
leaves us with 61 and 21 treated and control vessels, which have just
over 22 million observations where about 22\% are identified as fishing
events. For each of the remainig vessels, we calculate their total daily
fishing hours and obtain a pannel data with 38,200 observations. Table
\ref{tab:baci_n_s} shows the number of vessels following a BACI design,
as well as the fishing hours, before and after PIPA. Figure
\ref{fig:baci_strict} provides a visual representation of the
vessel-level fishing events that make up each group through time.

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-6}\label{tab:baci_n_s}Number of fishing vessels and mean daily fishing hours by group before and after PIPA implementation.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Group & n & Before & After & Change (A / B)\\
\midrule
Control & 26 & 11.74 & 20.13 & 1.72\\
Treatment & 61 & 10.67 & 18.36 & 1.72\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

The first analysis focuses on identifying the response of fishing
vessels to PIPA closure. Our variables of interests are daily fishing
hours, and daily distance traveled (Km). We compare our variable of
interest before and after the implementation of PIPA using a
Difference-in-Differences approach, where we track the variable of
interest for vessels that used to fish inside PIPA and vessels that
never fished inside PIPA, before and after PIPA implementation. Our
specification is the following:

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month-, and flag-level dummies that account for seasonality or
country-level management
interventions\footnote{We test for other more complex specifications that interact a quarterly or year-month variable with the treatment group and find qualitativel the same results.}

Our second part of the analyses focuses on the redistribution of fishing
effort. In other words, identifying where do vessels that used to fish
inside PIPA go after its establishment. We discretize spatial units by
using a polygon for PIPA
\footnote{we would expect to see a decrease here} and distinct spatial
units for each EEZ of each country. Some vessels might shift from EEZs
into the High Seas, but we are interested in knowing \emph{where} in the
High Seas, so we incorporate additional regions by using a 1 degree
buffer of the High Seas arround each of the EEZ regions. The rest of the
high seas are merged into a single spatial unit. For example, if we were
to do this only for Kiribati, we would have 8 spatial units: PIPA, three
EEZs, three 1-degree buffers of High Seas arround each EEZs, and the
rest of the High Seas. Whenever the buffers overlapped between
themselves, we randocmply clipped one on to the other. EEZs that had
sporadic fishing events were pooled into a group of ``others''.

To evaluate this change in effort allocation, we regress our variable of
interest (\emph{i.e.} fishing hours) on the interaction between a dummy
variable indicating the policy intervention and a dummy variable for
countries. This gives us the by-country change in proportional
allocation of fishing effort:

\[
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that country \(i\) receives at time \(t\). \(Post\) also
represents a policy dummy that takes the value of 0 for all dates before
implementation of PIPA, and 1 otherwise. \(Country\) is a dummy variable
for countries for the spatial units defined above. Our parameter of
interest is \(\beta_{1,i}\), which captures the country-level change in
proportional fishing effort.

All regression coefficients were estimated via ordinary least squares,
and heteroskedastic-robust standard errors were calculated. All analyses
were performed in R version 3.5.1 \citep{rcore_2018}. Raw data and code
used in this work are available on
\href{https://github.com/jcvdav/MPA_displacement}{github}.

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-9}\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.}
\centering
\begin{tabular}[t]{lr}
\toprule
country & change\\
\midrule
EEZ COK 1 & 0.53\\
EEZ FSM 1 & 0.81\\
EEZ KIR 1 & -0.16\\
EEZ KIR 2 & 4.50\\
EEZ KIR 3 & -2.79\\
\addlinespace
EEZ MHL 1 & -0.44\\
EEZ NRU 1 & 0.07\\
EEZ PNG 2 & -9.07\\
EEZ SLB 1 & 3.18\\
EEZ TUV 1 & 1.27\\
\addlinespace
HS & 3.93\\
HS COK 1 & 0.06\\
HS FSM 1 & 0.01\\
HS KIR 1 & 2.96\\
HS KIR 2 & 1.07\\
\addlinespace
HS KIR 3 & 2.53\\
HS MHL 1 & 0.11\\
HS MUS 1 & 0.00\\
HS NRU 1 & -0.08\\
HS PNG 2 & -0.20\\
\addlinespace
HS SLB 1 & -0.02\\
HS TUV 1 & 0.01\\
PIPA PIPA 1 & -8.25\\
\bottomrule
\end{tabular}
\end{table}

\clearpage

\hypertarget{results}{%
\section{Results}\label{results}}

Our data suggest that purse seiners and longliners have different
responses to the implementation of a Large-Scale Marine Protected Areas.
Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse
seiners have an abrupt increase, just before January 1st, 2015. This
trend is observed for both treated and control groups. The pattern
closely corresponds with the increase in total hours, but the total
number of vessels doesn't entirely follow this pattern. The increase in
fishing hours might be caused by the increased number of
satellites\footnote{Need to check this. Not sure any satellites were
  incorporated during 2014. It is also possible that PNA countries
  started enforcing the requirement of having an AIS unit. On either
  case, both treatment and control groups seem to be affected equally.}.
Longliners, however, show no apparent trend with a clear seasonality
\citep{ortuocrespo_2018}. The number of mmsi codes also increases
slightly through time, but becomes stable after 2015. For both gears and
across all measures, the treatment and control vessels follow similar
patterns, confirming our claim that the control group provides a
plausible counterfactual.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{\label{fig:unnamed-chunk-10}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed line
indicates PIPA closure.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{\label{fig:unnamed-chunk-11}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed line
indicates PIPA closure.}
\end{figure}

Our DiD analysis shows that treated purse seiners reduce their fishing
effort after PIPA implementation in the order of 16 hours per month.
This result is robust and significant (\emph{p \textless{} 0.05}) for
all model specifications, with the effect varying between
\(\beta_3 = -16.457\) and \(\beta_3 = -18.709\). Model specifications
that include the year polynomial show lower values for the \(\beta_1\)
coefficient associated to the \(Post_t\) policy dummy, and show positive
and negative values for \(\mu_1\) and \(\mu_2\), the linear and
quadratic terms for \(Y_t\), respectively. These effectively represent
the patterns observed in Figure \ref{fig:all_vessels}.

Regressions coefficients for each gear type are shown in Tables
\ref{tab:purse} and \ref{tab:long}. Column (1) presents the DiD
regression with no fixed effects, column (2) includes month fixed
effects, column (3) includes month and the second degree polynomial for
years, and column (4) includes all of the above and country-level fixed
effects.

Recall that to evaluate the redistribution of fishing effort we only
track fishing vessels that belong to the treated group. In this case, we
calculate the proportion of fishing effort allocated every month to each
spatially explicit region outlined by EEZs and the high seas. For purse
seiners, these represent 9 main EEZs, PIPA, the high seas, and a group
of other EEZs. Figure \ref{fig:redist_trend_ps} shows the monthly
relative fishing hours that each region received by all 176 treated
vessels. The top-left panel shows the change in fishing effort inside
PIPA, including the preemptive fishing and immediate reduction
previously reported \citep{mcdermott_2018}.

The change in the relative allocation of fishing effort by purse seiners
increases in eight of the 12 regions after PIPA implementation (Table
\ref{tab:disp_mod}). The largest increase is observed for the I-Kiribati
EEZ, with an average increase of 0.11 (\emph{p \textless{} 0.001}). In
other words, the redistribution of treated vessels caused a 10\%
increase in the \emph{relative} allocation of fishing effort within
I-Kiribati waters. The only decrease is observed for Papua New Guinea,
but the coefficient is not significant. Figure \ref{fig:map_change_ps}
provides a spatial representation of these changes. It is evident that
the increase in relative fishing effort is greater for for regions
closer to PIPA.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}
\caption{\label{fig:unnamed-chunk-15}\label{fig:redist_trend_ps}Monthly
relative allocation of fishing effort by PIPA-fishing vessels before and
after PIPA for 9 EEZs, PIPA, the high seas and `other EEZs'.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-16-1.pdf}
\caption{\label{fig:unnamed-chunk-16}\label{fig:map_change_ps}Spatial
representation of the mean change in the monthly allocation of fishing
effort for purse seiners.}
\end{figure}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-17}Change in the relative allocation of fishing hours by purse seiners ()}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{ll}
\toprule
term & h\_prop\\

\bottomrule
\end{tabular}
\end{table}

\clearpage

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Our findings provide interesting insights into the effect that LSMPAs
can have on vessel behavior and the redistribution of fishing effort.
These collection of results shows that the implementation of PIPA caused
treated vessels to reduce their fishing hours, and that this effect is
greater for purse seiners than longliners. Even though treated vessels
fish less, their relative allocation of fishing hours increased for all
other fishing grounds. This does not imply that there is more fishing
effort exerted by treated vessels, but rather that each region receives
a greater portion of the post-PIPA fishing effort of these same vessels,
which is lower than pre-PIPA levels. In this section we discuss the
implications of vessel-level reductions in fishing effort and the
increase in relative allocation of the remaining effort through space.
We also provide plausible explanations as to why purse seiners seem to
be more reactive to the spatial closure.

A major shortcoming of our analyses is that we do not observe catches or
revenues, which ultimately are the factors that guide the decision
making process of profit-maximizing agents. Therefore, it is difficult
to know whether the reduction in fishing effort represents a positive or
negative impact. A decrease in fishing effort is associated to an
increase in catches (and therefore greater CPUE) only when the entire
fleet does it, and if previous levels of effort were greater than
\(F_{MEY}\) (\emph{i.e.} the effort that would yield the maximum
economic yield). Therefore, it is plausible that the reduction of
fishing hours is not done by choice, but rather results from fishers
having to increase search time. Upon being relocated, fishers may not
identify the best fishing grounds as easily as before, and therefore
invest a greater proportion of their time searching for their catch.
Further analysis of temporal trends in non-fishing hours, as well as
distance traveled should provide us with insights as to why fishers
reduced fishing hours.

Previous studies on insular environments suggest that vessels move to
distant places, which might be translated as increased costs
\citep{stevenson_2013}. Nevertheless, they do not use counterfactuals
that could help account for system- or fleet-level changes that occur
through time. Others have used similar satellite-tracking systems to
show that fishing effort accumulates near the edges of spatial closures,
yielding greater catches \citep{murawski_2005}. Yet, these vessel tracks
do not cover the pre-reserve period, making it difficult identify the
contribution of spatial closures to the observed spatial distribution of
fishing vessels. Recent work by \citet{elahi_2018} identified that total
fishing effort in a focal region where a short-term MPA was implemented
showed little change, likely indicating that fishers redistributed
fishing effort to compensate for the reduction in available space. Our
data is assembled in a similar way, with fishing positions before and
after the implementation of PIPA and vessels grouped into treated and
control groups. Our BACI design, along with our
difference-in-differences analysis allows us to make causal inferences
about the effect that large scale marine protected areas have on fishing
effort.

The different responses observed between purse seiners and longliners
might have two possible explanations. It is likely that PIPA did not
contain habitat that longliners would consider optimal. Therefore, the
sporadic fishing events that occurred there are of little importance to
the fleet, and it is unlikely that the implementation of PIPA has an
effect on them. Alternatively, the differences may be due to the nature
of each fishing gear. Purse seiners are often constrained by seafloor
and termocline depth, and have a smaller spatial footprint
\citep{kroodsma_2018}. Tuna purse seiners are known to have greater
proportion of null sets (\emph{i.e.} where purse seines effectively cast
their nets, but no catch is obtained) during El Niño years, where the
termocline deepens in the Eastern Pacific \citep{dreyfusleon_2015}. On
the other hand, longliners may be more flexible as to where they can
deploy their longlines. \citet{ortuocrespo_2018} evaluated the
ecological niche of the pelagic longline fleet, and suggest that the
fleet may be under-utilizing the ocean, meaning that they can easily
redistribute elsewhere.

Our work suggests that the implementation of LSMPAs can have important
implications for purse seiners, and less so for longliners. We also show
that fishing effort is redistributed to areas close by. Future
management interventions that aim to close large portion of the oceans
should consider how fishing effort will change in space and through
time, and the ecological implications of this redistribution to ensure
that fishing effort is not just displaced elsewhere, leading to
overfishing in adjacent waters.

\clearpage

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
\caption{\label{fig:unnamed-chunk-18}\label{fig:baci_strict}Stream of
fishing events by vessels through time. Each line represents a vessel,
with dots indicating months with fishing activity and colors indicating
the pre and post periods.}
\end{figure}

\hypertarget{other-model-specifications}{%
\subsection{Other model
specifications}\label{other-model-specifications}}

\hypertarget{year-as-dummy-variables}{%
\subsubsection{Year as dummy variables}\label{year-as-dummy-variables}}

\begin{table}[!htbp] \centering 
  \caption{\label{tab:purse_old}Fishing hours from GFW for purse seiners (n = 106; 38 control, 68 treatment). Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robuste standard errors.} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{hours} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 post & 8.394$^{***}$ & 9.334$^{***}$ & 14.934$^{***}$ & 17.284$^{***}$ \\ 
  & (0.267) & (0.254) & (0.312) & (0.403) \\ 
  & & & & \\ 
 treated & $-$1.069$^{***}$ & $-$0.770$^{***}$ & $-$0.993$^{***}$ & 0.259 \\ 
  & (0.249) & (0.234) & (0.219) & (0.269) \\ 
  & & & & \\ 
 post:treated & $-$0.701$^{**}$ & $-$0.899$^{***}$ & $-$0.530$^{*}$ & $-$0.580$^{**}$ \\ 
  & (0.308) & (0.295) & (0.283) & (0.292) \\ 
  & & & & \\ 
 Constant & 11.738$^{***}$ & 11.281$^{***}$ & 7.574$^{***}$ & 10.077$^{***}$ \\ 
  & (0.220) & (0.292) & (0.318) & (0.422) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Month FE & No & Yes & Yes & Yes \\ 
Year FE & No & No & Yes & Yes \\ 
Flag FE & No & No & No & Yes \\ 
Observations & 37,840 & 37,840 & 37,840 & 37,840 \\ 
R$^{2}$ & 0.087 & 0.136 & 0.179 & 0.192 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\begin{table}[!htbp] \centering 
  \caption{\label{tab:long}Fishing hours from GFW for longliners (n = 203; 88 control, 115 treatment).. Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robuste standard errors.} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{hours} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 post & $-$1.241$^{***}$ & $-$1.026$^{***}$ & $-$0.944$^{***}$ & 13.292$^{***}$ \\ 
  & (0.116) & (0.116) & (0.157) & (0.330) \\ 
  & & & & \\ 
 treated & 2.595$^{***}$ & 2.708$^{***}$ & 2.639$^{***}$ & 1.493$^{***}$ \\ 
  & (0.101) & (0.101) & (0.102) & (0.116) \\ 
  & & & & \\ 
 post:treated & 0.053 & $-$0.069 & $-$0.029 & $-$0.060 \\ 
  & (0.133) & (0.133) & (0.134) & (0.135) \\ 
  & & & & \\ 
 Constant & 39.032$^{***}$ & 38.105$^{***}$ & 38.388$^{***}$ & 36.993$^{***}$ \\ 
  & (0.089) & (0.131) & (0.150) & (0.178) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Month FE & No & Yes & Yes & Yes \\ 
Year FE & No & No & Yes & Yes \\ 
Flag FE & No & No & No & Yes \\ 
Observations & 227,873 & 227,873 & 227,873 & 227,873 \\ 
R$^{2}$ & 0.010 & 0.015 & 0.016 & 0.030 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-22-1.pdf}
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-22-2.pdf}
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-22-3.pdf}
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-22-4.pdf}

\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-23-2.pdf}
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-23-3.pdf}
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-23-4.pdf}

\clearpage

\bibliography{references.bib}

\end{document}