### 2018-10-21 docs/Manuscript.tex
Updated all

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Displacement of fishing effort by Large Scale Marine Protected Areas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{chicago}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Displacement of fishing effort by Large Scale Marine Protected
Areas\footnote{Work in progress, do not circulate}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-20}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

\usepackage{float}
\floatplacement{figure}{H}
\usepackage{lineno}
\linenumbers
\usepackage[bottom]{footmisc}
\usepackage{pdflscape}
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\begin{document}
\maketitle
\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, but little empirical works
evaluate what happens to the effort after an MPA is created. The
economic and ecological implications of displacing fishing effort are
not yet fully understood. We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analyses of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large Scale Marine Protected Area in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) Vessels that fished inside the protected area
redistribute to adjacent waters; and 3) We observe a crowding effect for
the first years after implementation. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications. As countries continue to implement
LSMPAs as a way to reach the stated 10\% target of ocean protection,
managers should consider how fishing effort will change in space and
through time to ensure that fishing effort is not just displaced
elsewhere, leading to overfishing in adjacent waters. While LSMPAs can
provide a wide range of benefits, their implementation must be
accompained with traditional fisheries management to maximize their
effectiveness.
\end{abstract}

\clearpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environments by
2020. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by
a small number of Large Scale Marine Protected Areas (LSMPA)
\citet{singleton_2014,boonzaier_2016,alger_2017}). Due to weak property
rights, limited habitat transformation, and potentially lower management
costs, LSMPAs provide an opportunity to safeguard the oceans
\citet{game_2009}. Today, a small number of LSMPAs make up at least 80\%
of the managed areas in the ocean \citet{toonen_2013}. LSMPAs may be
instrumental in helping us achieve the 10\% target by 2020
\citep{mccauley_2016}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. As with customary MPAs, it is important that we
understand the socioeconomic implications of management interventions.
One issue of particular importance is that of the displacement or
redistribution of fishing effort, which may influence the outcomes of an
MPA \citep{smith_2003}. Theoretical models make different assumptions
about the way in which fishers will reallocate fishing effort after an
area closure. The few empirical works focus on small-scale fisheries and
rarely incorporate spatially explicit data. Despite these efforts,
marine conservation is yet to identify the response in fishing effort
and its management implications.

Recent advances in satellite tracking technologies and near real-time
identification of fishing activity provide us with an opportunity to
tackle increase governance and answer long-lasting questions. Here, we
ask, how do fishers respond to the implementation of LSMPA? After a
closure, and if they continue to fish, where do they reallocate fishing
effort? We use identification of fishing activity via Automatic
Identification Systems (AIS) and causal inference techniques to describe
the behavioral changes and spatial redistribution of the industrial tuna
purse seine fleet due to the implementation of a Large Scale Marine
Protected Areas in the Pacific Ocean.

Our work is novel in the sense that it provides empirical evidence of
the effect of Large Scale Marine Protected Areas in fishing behavior and
distribution and can help guide future interventions as countries march
towards meeting the targeted 10\% of ocean protection. Understanding how
effort is displaced from LSMPAs might provide insights into how the
industrial fleet would react to a high seas closure.

The paper is outlined as follows: Section \ref{background} provides more
information on Large Scale Marine Protected Areas and background on
empirical studies of effort redistribution. An overview of the Nauru
Agreement and associated countries, a description of the fleet that
operates in the region, and a brief history of PIPA is also included.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, section \ref{discussion}
provides an extension of our results to other
LSMPAs\footnote{Not yet. but I think this would be interesting} and
discusses our
results\footnote{I might also include a section, after the background, where I explain the different theoretical models of effort redistribution and behavioral responses}.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{large-scale-marine-protected-areas}{%
\subsection{Large Scale Marine Protected
Areas}\label{large-scale-marine-protected-areas}}

The cutoff at which an MPA is considered to be a LSMPA ranges from areas
larger than 30,000 km\textsuperscript{2} as defined by
\citet{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
LSMPAs that meet the later, and that are fully no take. LSMPAs are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The
early literature on LSMPAs focused on the inherent challenges and
difficulties that come with a pelagic environment. \citet{kaplan_2010}
claimed that very large MPAs would result in excessive opportunity costs
and that these would be difficult to enforce. \citet{game_2009}
suggested that most of the challenges could be overcome with the
incorporation of technology, in what then became known as Dynamic Ocean
Management \citep{maxwell_2015}. The effectiveness and difficulties of
implementing LSMPAs are not the focus of this paper
\footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}.
Instead, we focus on the effect and implications of the ones that
\emph{have been} implemented.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{\label{fig:unnamed-chunk-13}\label{fig:LSMPAs_map}Large Scale
Marine Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated english name is `Protected Area'.}
\end{figure}

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protects two
thirds of the tagged grey reef sharks by effectively excluding fishing
effort. More recently, \citep{bradley_2018} use similar data to
highlight cases of potential illegal shark fishing \emph{inside} a 2
million km\textsuperscript{2} shark sanctuary. To date, no studies have
evaluated the displacement of fishing effort due to LSMPA
implementation.

Spatial closures of this magnitude are likely to induce changes in
fishers' behavior. Theoretical models of fishing effort redistribution
range from the simplistic assumption that effort inside the bounded
region disappears, to spatially explicit models that reallocate fishing
effort based on habitat characteristics, presence of other vessels, and
expected returns \citep{smith_2003,hilborn_2006}. However, these focus
on the long term optimal equilibrium, and redistribution of fishing
effort may not always be optimal, especially over the first years
\citep{stevenson_2013}.

The empirical research that has been done in customary sized MPAs
suggest that resource users may show idiosyncratic responses. For
example, \citet{stevenson_2013} show that a network of MPAs displaced
fishing effort farther away from ports, resulting in higher
\emph{perceived} costs, and increases in catch per unit effort.
\citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of a network of MPAs in
California, and find that commercial dive boats follow a
fishing-the-line pattern, while some fishing boats follow an ideal free
distribution. More recently \citet{elahi_2018} used satellite tracking
data to show that a temporal spatial closure caused trawlers to maintain
effort but apply it more intensively elsewhere, particularly along the
borders and closer to shore. The way in which fishers react to a spatial
closure can have major implications in its outcome
\citep{smith_2003,hilborn_2006} highlighting the need to understand how
fishers react to the implementation of LSMPAs, and fishing effort
changes and is spatially redistributed. However, all these closures took
place within Exclusive Economic Zones, where other regulations exist.
This may not always be the case for LSMPAs, where often the entire EEZ
is converted into a LSMPA, leaving fishers with the options of moving to
the high seas or other countries' EEZs.

\hypertarget{nauru-agreement-and-the-phoenix-island-protected-area}{%
\subsection{Nauru agreement and the Phoenix Island Protected
Area}\label{nauru-agreement-and-the-phoenix-island-protected-area}}

The Nauru Agreement was established in 1982 by a select group of Pacific
island nations to manage their important tuna resources. PNA Members
include Federated States of Micronesia, Kiribati, Marshall Islands,
Nauru, Palau, Papua New Guinea, Solomon Islands, and Tuvalu. The Nauru
Agreement regulated access of foreign vessels (\emph{i.e.} those from
non-PNA countries). Holding \textasciitilde{}80\% of the historical
purse seining grounds within their Exclusive Economic Zones, PNA
countries gained bargaining power when providing access to foreign
fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisers recommend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then use or sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}
\caption{\label{fig:unnamed-chunk-14}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. Countries that belong
to the PNA are shown in blue, while empty polygons indicates all others.
A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. This region of the
Pacific has historically accounted for a large portion of tuna catches
\citep{aqorau_1997}. Today, the PNA controls close to 50\% of the global
skipjack tuna production \citep{pna_website_2018}. A large portion of
these catches derive from purse seine vessels licensed under the VDS.
Fishing vessels from Australia, New Zealand, China, France, Korea,
Japan, the Philippines, Taiwan, and the United States participate in the
purse-seining VDS.

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Island Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of it was declared as no-take. In
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. Figure \ref{fig:PNA_map} shows a map of the
PNA countries and the Phoenix Island Protected Area.

The closure of such a large area in one of the most important fishing
regions in the world provides a great opportunity to evaluate the
behavioral responses and redistribution of fishing effort by vessels
that used to fish there. PIPA has been the focus of previous research,
showing that fishing effort is effectively reduced after implementation
\citep{mccauley_2016,mcdermott_2018}. To this, we pose two questions:
How do individual vessels respond to the sudden exclusion of such a big
area? And where did all the vessels go? In the next sections we describe
the data and methods used to answer this questions.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

This section is divided into two main parts. First, we provide a general
description of AIS data and the process of identification of
vessel-level fishing events done by Global Fishing Watch\footnote{Global
  Fishing Watch: \url{globalfishingwatch.org}}. Alongside, we describe
the subset of data used on our analyses. When relevant, we also point
out possible shortcomings in the data, or factors that must be
considered in the later analyses. We then move on to explain our
empirical strategy for the identification of the behavioral changes and
redistribution of fishing effort.

\hypertarget{data}{%
\subsection{Data}\label{data}}

Automatic Identification Systems (AIS) are on-board devices that provide
at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activity to surrounding vessels. These broadcasted
messages can be received by satellites and land-based antennas. GFW uses
convolutional neural networks to infer vessel characteristics and
whether each broadcasted position represents a fishing event, thus
allowing us to estimate near real-time fishing events globally since
2012 \citep{kroodsma_2018}.

The amount of data gathered is dependent on the number of antennas and
satellites that can pick up the signals. In June 1\textsuperscript{st}
2014 satellite count increased from 3 to 6, and then to 10 in January
1\textsuperscript{st} 2016. This causes an increase in the number of
\emph{received} AIS messages (\emph{i.e.} points), and therefore an
apparent increase in the number of vessels and fishing hours. However,
the addition of satellites ``affects'' all vessels in the same way. The
variability in AIS data and ocean conditions require that temporal
trends be taken into account. We do that by including specific controls
in our identification strategy and using a subset of data that meet a
BACI design -which gives us the full tracks for vessels affected and
unaffected by the implementation of PIPA.

Our analysis focuses on purse seine vessels, the most important fishery
for PNA
countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}.
We identify a total of 103 purse seiners that fished in PNA waters at
least once before
2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses.}.
These vessels represent over 26 million individual observations for the
2012 - 2017 period. We identify 65 vessels that have fished inside PIPA
at least once since 2012. From these, 62 did so before the announcement
(\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}) but are
only able to track 61 for the complete period of study before and after
its
implementation\footnote{1 vessel never fished again after August 18, 2013.}.
On the other hand, 38 vessels never fished inside PIPA but we only have
data for 26 vessels before and after.

Therefore, our treatment group contains all purse seiners (n = 61) that
fished within PIPA at least once before the announcement, and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure (n = 26).

We include three additional definitions of control groups as a
robustness check: one with only vessels that belong to PNA countries (n
= 7), and one that excludes Chinese vessels (n = 21). Our third control
is made up of Japanese purse seiners that fish in the Pacific but that
never fished inside PIPA (n =
27)\footnote{Using Taiwanese vessels was the original idea, but there are only 4 vessels with pre-2015 data. I need to further identify proper Taiwanese vessels, as I am currently using the entire Pacific ocean}.
Our definition of treatment and control groups leaves us with 61 and 21
treated and control vessels, which have just over 22 million
observations where about 22\% are identified as fishing events.

For each vessel we calculate total daily fishing hours and obtain panel
data with 37,800 observations. Table \ref{tab:baci_n_s} shows the number
of vessels following a BACI design, as well as the fishing hours, before
and after PIPA. Fig. \ref{fig:all_vessels} shows that mean fishing hours
for purse seiners have an abrupt increase, just before January 1st,
2015. This trend is observed for both treated and control groups. Across
all measures, the treatment and control vessels follow similar patterns,
confirming our claim that the control group provides a plausible
counterfactual. Figure \ref{fig:baci_strict} provides a visual
representation of the vessel-level fishing events that make up each
group through time. We use this data to answer our key questions: How do
the 61 purse seiners modify their behavior as compared to the different
control groups, and where do they go after the spatial closure? The
following section describes our empirical identification strategies.

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-17}\label{tab:baci_n_s}Number of fishing vessels and mean daily fishing hours by group before and after PIPA implementation.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Group & n & Before & After & Change (A / B)\\
\midrule
Control & 26 & 11.74 & 20.13 & 1.72\\
Treatment & 61 & 10.67 & 18.36 & 1.72\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
\caption{\label{fig:unnamed-chunk-18}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed lines
indicates dates when satellites were added, solid line indicates PIPA
closurePIPA closure.}
\end{figure}

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

The first analysis focuses on identifying the response of fishing
vessels to PIPA closure. A spatial closure might cause fishers to modify
their behavior as they adapt to a new state of the world. Some may have
to travel further distances to find new fishing grounds, which would
lead them to fish harder to compensate for the associated extra fuel
costs. If fishers had developed experience for fishing in particular
sites, being excluded might impose a learning cost on them, as they
identify new fishing grounds. This would result in increased search
times. However, if the area of influence of a vessel is significantly
larger than that of the spatial closure, they might already know other
places and times where to fish. Our variables of interests intend to
capture these possible reactions.

We use daily fishing hours, daily proportion of fishing vs.~non-fishing
hours, and daily distance traveled (km). We compare our variable of
interest before and after the implementation of PIPA using a
Difference-in-Differences approach, where we track the variable of
interest for vessels described in the previous section. Our
specification is the following:

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \theta_t + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month- and flag-level dummies that account for seasonality or
country-level management interventions. \(\theta_t\) are dummies to
account for the addition of more satellites.
\footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}.

Our second part of the analyses focuses on the redistribution of fishing
effort. The role of institutions may play an important role. As stated
before, LSMPAs often span the entirety of an EEZ. In the case of purse
seining in the PNA, vessels purchase access to the fishery at the
beginning of the season. If a vessel decides to fish within the PNA,
they've already made the decision to pay for access, expecting higher
returns from fishing there than in the High Seas. In the particular case
of PIPA, one would expect that a vessel holding a permit to fish in
I-Kiribati waters would then move to other fishing grounds within
Kiribati after the closure. Alternatively, if a vessel fishes illegally
in I-Kiribati waters before the implementation, one would expect them to
reallocate to other regions with equal or less enforcement. In this
case, they might then choose to move to other I-Kiribati waters, waters
of countries that have similar enforcement levels, or the High Seas.

We discretize spatial units by using a polygon for
PIPA\footnote{we would expect to see a decrease here} and distinct
spatial units for each EEZ of each country. Some vessels might shift
from EEZs into the high seas, but we are interested in knowing
\emph{where} in the high seas, so we incorporate additional regions by
using a 1 degree buffer of the high seas around each of the EEZ regions.
The rest of the high seas are merged into a single spatial unit. For
example, if we were to do this only for Kiribati, we would have 8
spatial units: PIPA, three EEZs, three 1-degree buffers of high seas
around each EEZs, and the rest of the high seas. Whenever the buffers
overlapped between themselves, we randomly clipped one on to the other.

We then take this spatial polygons and rasterize them to a 1-degree grid
(See Figure \ref{fig:raster_rgn}), which we then use to rasterize points
of fishing activity. For each cell, we calculate the total monthly
fishing hours by treated and control vessels (See Figure
\ref{fig:fishing_raster}). Since we are interested in the redistribution
of fishing effort, we use this gridded data to calculate the proportion
of fishing hours that each region receives each month relative to the
total fishing hours observed across all regions and all cells in that
month.

We use this measure to evaluate the change in effort allocation by
regressing it on the interaction between a year dummy and a dummy for
regions:

\[
y_{i,t} = \alpha + \beta_1Year_t + \beta_{2,i}Region_i + \beta_{3,i}Year_t \times Region_i+ \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that region \(i\) receives at time \(t\). Years are
modelled as factors (\(Year_t\)), using 2012 as the reference level.
\(Region\) is a dummy variable for regions defined above. Our parameter
of interest is \(\beta_{3,i}\), which captures the yearly by-country
change in proportional allocation of fishing effort relative to 2012.

All regression coefficients were estimated via ordinary least squares,
and heteroskedastic-robust standard errors were calculated. All analyses
were performed in R version 3.5.1 \citep{rcore_2018}. Raw data and code
used in this work are available on
\href{https://github.com/jcvdav/MPA_displacement}{github}.

\clearpage

\hypertarget{results}{%
\section{Results}\label{results}}

Regressions coefficients for our DiD analysis are shown in Table
\ref{tab:did}. Columns 1 -3 use all data for controls. Columns 4 - 6 use
only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels.
Columns 10 and 11 use Japanese vessels in the Pacific as controls, and
hence don't include flag fixed-effects to avoid perfect colinearity
between Japanese flag and
control\footnote{Results of the same analysis is shown for longliners in \ref{tab:long}}.
Our DiD analysis shows an overall increase in purse seine fishing hours,
even after accounting for the introduction of new satellites (Table
\ref{tab:did}). This coefficient estimate is consistent for different
model specifications and across groups of treatment and controls, and
effectively represents the patterns observed in Figure
\ref{fig:all_vessels}.

The \(\beta_3\) coefficient indicating our treatment effect suggests
that, relative to the control, treated vessels fish less, in the order
of 0.5 hours per day. Another way to interpret this is that the increase
in fishing effort by treated vessels has occurred at a lower rate than
control vessels. The effect varying between \(\beta_3 = -0.515\) and
\(\beta_3 = -0.934\). These values are equivalent to a 15 - 28 reduction
in fishing hours per month. However, this result is only robust and
significant (\(p < 0.01\)) in the simplest specification of the
Diff-in-Diff. When reducing the linear structure of the
\(Pre \times Post\) design and instead interacting the treatment dummy
with quarterly or year-month combinations we are not able to reject the
null hypotheses of no change, suggesting that the total fishing hours
remains similar for the displaced vessels (Figs.
\ref{fig:q1}-\ref{fig:ym4}). Since vessels don't seem to be fishing less
(or at least not much less), an obvious question arises: where are they
fishing now?

Figure \ref{fig:fishing_raster} provides a spatial representation of
these changes. Note that 2013 effort distribution is roughly similar
across vessels. Then in 2014 there is a sharp increase in fishing hours
by treated vessels inside PIPA (blue paradox). In 2015 treated vessels
then allocate more effort to the easternmost Kiribati EEZ (see Figure
\label{fig:PNA_map}) and the high seas. In 2016 spatial distribution of
fishing effort is again similar across groups. It is evident that the
increase in relative fishing effort is greater for for regions closer to
PIPA.

Along a similar line, Figure \ref{fig:redist_trend} shows a detailed
temporal evolution of the relative allocation of fishing
effort\footnote{Recall that to evaluate the redistribution of fishing effort we only track fishing vessels that belong to the treated group.}.
It shows shows the change in fishing effort inside PIPA, including the
preemptive fishing and immediate reduction previously reported
\citep{mcdermott_2018}. I-Kiribati waters in KIR 1 and KIR 2 show an
increase in fishing effort for 2015; the same pattern is observed for
their respective High Seas buffers, and the general High Seas. Other
regions farther away from PIPA don't show a clear pattern under visual
inspection.

We then move on to our regression results, presented as a figure showing
the year-region interaction coefficient estimates for each region as a
stream of changes through time (Fig. \ref{fig:mean_change} and Table
\ref{tab:mean_change}). As noted before, the marginal change in the
relative allocation of fishing effort relative to 2012 increases in 2014
for PIPA. Most coefficients are not statistically significant, but
follow the same trends previously discussed, with a tendency for regions
closer to PIPA to show an increase in the relative allocation of fishing
effort in the post-implementation period. The largest increase is
observed for the I-Kiribati EEZ 2, where the redistribution of treated
vessels caused a 15\% increase in the \emph{relative} allocation of
fishing effort within I-Kiribati waters. This displacement is likely
causing a crowding effect after PIPA-fishing vessels redistribute to
other areas.

We use the rasterized effort to evaluate if the PIPA closure increased
crowdness otuside the LSMPA. We count the number of raster cells that
had both treated and control vessels each year (Fig.
\ref{fig:sp_corr}A). The pattern observed here could just be an artifact
of satellite detections increasing though time. Thereforem we also
calculate a spatial correlation of presence/absence and observe similar
patterns (Fig. \ref{fig:sp_corr}B). While this are at the year level and
cannot assure that vessels were in effect at the same time in the same
vessel, it provides an accurate depiction of what we would expect to
find if our claims of displacement are true. Spatial overlap increases
in 2015 as vessels from PIPA are excluded, and that measure slowly
returns to its previous state.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-21-1.pdf}
\caption{\label{fig:unnamed-chunk-21}\label{fig:all_vessels}Daily fishing
hours for all vessels in our main treatment-control groups. Solid
straight lines show a linear trend by period (pre-psot) and treatment.
The other red and blue lines show monthly avarages. Vertical dashed
lines indicates dates when satellites were added, solid line indicates
PIPA closurePIPA closure.}
\end{figure}

\clearpage
\begin{landscape}

\begin{table}[!htbp] \centering 
  \caption{\label{tab:did}Difference-in-differences estimates for 3 different controls and 3 different specifications. The first three columns use all data for controls. Columns 4 - 6 use only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels. Columns 10 and 11 use Japanese vessels in the Pacific as controls, and hence don't include flag FEs. Numbers in parentheses are heteroskedastic-robust standard errors.} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{1pt}}lccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{11}{c}{\textit{Dependent variable:}} \\ 
\cline{2-12} 
\\[-1.8ex] & \multicolumn{11}{c}{hours} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11)\\ 
\hline \\[-1.8ex] 
 Constant & 6.345$^{***}$ & 7.863$^{***}$ & 10.322$^{***}$ & 7.124$^{***}$ & 9.007$^{***}$ & 9.933$^{***}$ & 6.057$^{***}$ & 7.584$^{***}$ & 10.194$^{***}$ & 22.625$^{***}$ & 24.731$^{***}$ \\ 
  & (0.186) & (0.274) & (0.404) & (0.265) & (0.341) & (0.620) & (0.193) & (0.279) & (0.538) & (0.629) & (0.683) \\ 
  & & & & & & & & & & & \\ 
 post & $-$0.052 & 1.134$^{***}$ & 1.115$^{***}$ & $-$1.199$^{***}$ & $-$0.279 & 0.124 & $-$0.284 & 1.013$^{***}$ & 1.166$^{***}$ & 2.330$^{***}$ & 3.991$^{***}$ \\ 
  & (0.303) & (0.304) & (0.339) & (0.435) & (0.429) & (0.436) & (0.318) & (0.318) & (0.357) & (0.890) & (0.877) \\ 
  & & & & & & & & & & & \\ 
 treated & $-$1.142$^{***}$ & $-$0.839$^{***}$ & 0.010 & $-$1.977$^{***}$ & $-$1.675$^{***}$ & 0.626 & $-$0.877$^{***}$ & $-$0.536$^{**}$ & 0.150 & $-$18.190$^{***}$ & $-$18.257$^{***}$ \\ 
  & (0.214) & (0.209) & (0.281) & (0.284) & (0.274) & (0.416) & (0.221) & (0.215) & (0.297) & (0.673) & (0.661) \\ 
  & & & & & & & & & & & \\ 
 sate2 & 12.579$^{***}$ & 11.589$^{***}$ & 11.340$^{***}$ & 12.709$^{***}$ & 11.850$^{***}$ & 11.749$^{***}$ & 12.631$^{***}$ & 11.599$^{***}$ & 11.349$^{***}$ & 14.346$^{***}$ & 12.967$^{***}$ \\ 
  & (0.197) & (0.199) & (0.232) & (0.212) & (0.213) & (0.246) & (0.199) & (0.201) & (0.235) & (0.303) & (0.309) \\ 
  & & & & & & & & & & & \\ 
 sate3 & 14.675$^{***}$ & 13.587$^{***}$ & 13.328$^{***}$ & 14.799$^{***}$ & 13.894$^{***}$ & 13.795$^{***}$ & 14.958$^{***}$ & 13.804$^{***}$ & 13.566$^{***}$ & 15.187$^{***}$ & 13.719$^{***}$ \\ 
  & (0.260) & (0.262) & (0.307) & (0.286) & (0.287) & (0.329) & (0.264) & (0.266) & (0.315) & (0.402) & (0.410) \\ 
  & & & & & & & & & & & \\ 
 post:treated & $-$0.515$^{*}$ & $-$0.833$^{***}$ & $-$0.934$^{***}$ & 0.562 & 0.373 & $-$0.176 & $-$0.439 & $-$0.814$^{***}$ & $-$1.084$^{***}$ & $-$3.209$^{***}$ & $-$3.715$^{***}$ \\ 
  & (0.281) & (0.276) & (0.310) & (0.413) & (0.402) & (0.407) & (0.293) & (0.287) & (0.326) & (0.811) & (0.799) \\ 
  & & & & & & & & & & & \\ 
\hline \\[-1.8ex] 
Control & All & All & All & PNA & PNA & PNA & -CHN & -CHN & -CHN & JPN & JPN \\ 
Month FE & No & Yes & Yes & No & Yes & Yes & No & Yes & Yes & No & Yes \\ 
Flag FE & No & No & Yes & No & No & Yes & No & No & Yes & No & No \\ 
Observations & 37,840 & 37,840 & 30,359 & 30,583 & 30,583 & 25,034 & 36,415 & 36,415 & 28,934 & 34,047 & 34,047 \\ 
R$^{2}$ & 0.171 & 0.200 & 0.208 & 0.178 & 0.208 & 0.215 & 0.173 & 0.203 & 0.211 & 0.260 & 0.280 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{11}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\end{landscape}
\clearpage

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}
\caption{\label{fig:unnamed-chunk-23}\label{fig:long}Coefficient estimates
for each model. Top pannel indicates variable, x-axis represents model
specification, and y-axis coefficient estimate.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-24-1.pdf}
\caption{\label{fig:unnamed-chunk-24}\label{fig:fishing_raster}Yearly
spatial distribution of fishing effort by treated and control vessels.
Colors have been adjusted relative to the maximum observed by group and
year. Red polygons show LGMPAs in the region. For visualization purpose
sonly some years are presented, the full 2012 - 2017 period is shown in
Figure \ref{fig:fishing_raster_full}.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-25-1.pdf}
\caption{\label{fig:unnamed-chunk-25}\label{fig:redist_trend}Monthly
relative allocation of fishing effort by PIPA-fishing vessels before and
after PIPA by region. Colors indicate the pre-post period. Solid line
across points shows local mean with a region-specific loess smoother.
Vertical dashed lines indicates dates when satellites were added, solid
line indicates PIPA closurePIPA closure. Note that inflection points of
the loess smoother are not affected by the addition of satellites.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-26-1.pdf}
\caption{\label{fig:unnamed-chunk-26}\label{fig:mean_change}Coefficient
estimates for the redistribution regression. Each pannel shows the
region-specific coefficients, with heteroskedastick-robust standard
errors as error bars. The horizontal dashed line represents 0 chenge
relative to the 2012 region-specific levels.}
\end{figure}

\begin{table}

\caption{\label{tab:unnamed-chunk-27}\label{tab:mean_change}Coefficient estimates for the interaction of year and region. Each row represents a region, each color a year. Numbers in parentheses are heteroskedatic-robust standard errors. R2 = 0.47 (F(95, 1056) = 9.99; p < 0.001). * < 0.1; ** p < 0.05; ***p < 0.01.}
\centering
\begin{tabular}[t]{llllll}
\toprule
term & 2013 & 2014 & 2015 & 2016 & 2017\\
\midrule
PIPA PIPA 1 & -0.062 (0.026)** & 0.085 (0.043)** & -0.086 (0.026)*** & -0.079 (0.025)*** & -0.074 (0.025)***\\
EEZ KIR 1 & -0.034 (0.030) & 0.035 (0.038) & 0.042 (0.045) & -0.035 (0.031) & -0.023 (0.032)\\
EEZ KIR 2 & 0.001 (0.009) & 0.032 (0.019)* & 0.142 (0.061)** & 0.014 (0.013) & 0.014 (0.015)\\
EEZ KIR 3 & -0.188 (0.075)** & -0.061 (0.072) & -0.118 (0.069)* & -0.093 (0.083) & -0.098 (0.083)\\
HS KIR 1 & -0.003 (0.011) & 0.016 (0.012) & 0.038 (0.013)*** & 0.004 (0.013) & 0.037 (0.018)**\\
\addlinespace
HS KIR 2 & 0.009 (0.006) & 0.004 (0.006) & -0.000 (0.009) & 0.022 (0.010)** & 0.008 (0.006)\\
HS KIR 3 & -0.002 (0.009) & 0.003 (0.009) & -0.008 (0.011) & 0.022 (0.011)* & 0.064 (0.024)***\\
HS HS 1 & 0.021 (0.015) & 0.027 (0.017) & 0.045 (0.017)*** & 0.030 (0.017)* & 0.014 (0.015)\\
EEZ FSM 1 & 0.018 (0.033) & -0.015 (0.025) & 0.000 (0.037) & 0.012 (0.025) & -0.001 (0.021)\\
EEZ MHL 1 & 0.037 (0.022) & 0.025 (0.010)** & -0.006 (0.009) & 0.052 (0.032) & -0.000 (0.006)\\
\addlinespace
EEZ NRU 1 & 0.046 (0.021)** & 0.044 (0.020)** & 0.003 (0.017) & 0.033 (0.015)** & 0.050 (0.025)**\\
EEZ PNG 2 & 0.196 (0.072)*** & -0.106 (0.041)*** & -0.132 (0.043)*** & -0.064 (0.048) & 0.001 (0.049)\\
EEZ SLB 1 & 0.083 (0.056) & -0.019 (0.032) & -0.009 (0.038) & 0.081 (0.053) & 0.076 (0.046)*\\
EEZ TUV 1 & -0.055 (0.037) & -0.033 (0.040) & -0.063 (0.037)* & 0.018 (0.046) & -0.026 (0.038)\\
HS COK 1 & -0.003 (0.008) & -0.004 (0.008) & -0.003 (0.011) & -0.007 (0.008) & -0.003 (0.008)\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-28-1.pdf}
\caption{\label{fig:unnamed-chunk-28}\label{fig:sp_corr}Number of cells that
had treated and control vessels (A) and spatial correlation in the
presence-absence of treated and control vessels per cell (B).}
\end{figure}

\clearpage

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Our findings provide interesting insights into the effect that LSMPAs
can have on vessel behavior and the redistribution of fishing effort.
These collection of results show that the implementation of PIPA had
little effect on the effort exerted by purse seiners. However, regions
adjacent to PIPA show an increase in the relative amount of fishing
effort, suggesting that effort is simply redisplaced to closeby areas.
This does not imply that there is more fishing effort exerted by treated
vessels, but rather that each region receives a greater portion of the
post-PIPA fishing effort of these same vessels, which is lower than
pre-PIPA levels. Additionally, we show that post-PIPA there is a
concentration of effort leading to crowding. In this section we discuss
the implications of vessel-level reductions in fishing effort and the
increase in relative allocation of the remaining effort through space.
We also provide plausible explanations as to why purse seiners seem to
be more reactive to the spatial closure.

Previous studies on insular environments suggest that vessels move to
distant places, which might be translated as increased costs
\citep{stevenson_2013}. Nevertheless, they do not use counterfactuals
that could help account for system- or fleet-level changes that occur
through time. Others have used similar satellite-tracking systems to
show that fishing effort accumulates near the edges of spatial closures,
yielding greater catches \citep{murawski_2005}. Yet, these vessel tracks
do not cover the pre-reserve period, making it difficult identify the
contribution of spatial closures to the observed spatial distribution of
fishing vessels. Recent work by \citet{elahi_2018} identified that total
fishing effort in a focal region where a short-term MPA was implemented
showed little change, likely indicating that fishers redistributed
fishing effort to compensate for the reduction in available space. Our
data, which is assembled in a similar way, allows us to make similar
inferences about the unobserved change in aggregate fishing effort and
its spatial redistribution.

The role of the institutions at play is evident. Vessels that fished
inside PIPA before the implementation were, ultimately, fishing in
I-Kiribati waters likely under a VDS license. Uppon the implementation,
some of these redistribute to other areas of the I-Kiribat EEZ and the
high seas for the year immediately after the closure. As stated before,
it is reasonable for a vessel holding a license to fish in I-Kiribati
waters to simply reallocate their effor to other parts of that EEZ.
However, the shift to the high seas may not follow the same logic.
Instead, it is possible that vessels that redistribute to the high seas
were opperating in I-Kiribati illegally, and decide to reallocate in the
high seas to continue fishing without a license. These patterns provide
us with interesting insight sinto fishing behavior, which may become
useful in tackling Illegal, Unreported and Unregulted (IUU) fisheries.

A major shortcoming of our analyses is that we do not observe catches or
revenues, which ultimately are the factors that guide the decision
making process of profit-maximizing agents. Therefore, it is difficult
to know whether lack of change in fishing hours represents a positive or
negative impact. An additional thing that we are yet to test is the
changes in non-fishing hours sepnt at sea. It is plausible that fishing
hours remain constant, but vessels have to increase search time. Upon
being relocated, fishers may not identify the best fishing grounds as
easily as before, and therefore invest a greater proportion of their
time searching for their catch. Further analysis of temporal trends in
non-fishing hours, as well as distance traveled should provide us with
insights as to why fishers reduced fishing hours. However, the
widespread footprint of the treated fleet may suggest that they are well
aqquainted with the region (See Fig. \ref{fig:fishing_raster}), and its
unlikely that they would need to learn or identify new fishing grounds.

Our work suggests that the implementation of LSMPAs has little impact on
purse seiners fishng effort. We also show that fishing effort is
redistributed to areas close by, and that this leads to a potentially
negative crowding effect. A growing body of literature suggests that
closing the high seas to all fishing could increase fishery yields and
profitability of fisheries, with negligible costs to food security
\citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}. Such
management interventions should consider how fishing effort will change
in space and through time, and the ecological implications of this
redistribution to ensure that fishing effort is not just displaced
elsewhere, leading to overfishing in adjacent waters. While LSMPAs can
provide a wide range of benefits, their implementation must be
accompained with traditional fisheries management to maximize
effectiveness.

\hypertarget{further-work}{%
\section{Further work}\label{further-work}}

\begin{itemize}
\tightlist
\item
  Distance traveled
\item
  Non-fishing hours at sea
\item
  Monthly crowdnes??
\item
  Repeat raster exercise for Revillagigedo and other recently
  implemented LSMPAs. It would be great to have a pannel figure of 3 or
  4 other LSMPAs that show wher effort was and where it went to
  generalize our results.
\item
  John talked about this being relevant to Palau (their upcoming LSMPA).
  We might want to talk about where the vessels that are there would go.
\end{itemize}

\bibliography{references.bib}

\clearpage

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\hypertarget{stream-of-data-by-vessel-group-and-period}{%
\subsection{Stream of data by vessel, group, and
period}\label{stream-of-data-by-vessel-group-and-period}}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-29-1.pdf}
\caption{\label{fig:unnamed-chunk-29}\label{fig:baci_strict}Stream of
fishing events by vessels through time. Each line represents a vessel,
with dots indicating months with fishing activity and colors indicating
the pre and post periods.}
\end{figure}

\clearpage

\hypertarget{longliners}{%
\subsection{Longliners}\label{longliners}}

Longliners show a similar pattern of effort reduction. However, the
magnitude of the \(\beta_3\) coefficient is smaller (ranging from
\(\beta_3 = -0.98\) to \(\beta_3 = -1.125\)) and not significant unless
including month and flag FEs for the general treatment-control groups,
or when excluding Chinese vessels. This, along with higher standard
error values suggest that longliners have a smaller and more variable
response to the implementation of LSMPAs.

\begin{table}[!htbp] \centering 
  \caption{\label{tab:long}Difference-in-differences estimates for 2 different controls and 3 different specifications. The first three columns use all data for controls. Columns 4 - 6 exclude Chinese vessels. Numbers in parentheses are heteroskedastic-robust standard errors.} 
  \label{} 
\small 
\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{hours} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 Constant & 39.192$^{***}$ & 38.411$^{***}$ & 36.520$^{***}$ & 39.402$^{***}$ & 38.636$^{***}$ & 34.402$^{***}$ \\ 
  & (0.095) & (0.132) & (0.169) & (0.102) & (0.140) & (0.236) \\ 
  & & & & & & \\ 
 post & $-$1.014$^{***}$ & $-$0.189 & 0.710$^{***}$ & $-$1.400$^{***}$ & $-$0.574$^{***}$ & 0.896$^{***}$ \\ 
  & (0.143) & (0.145) & (0.158) & (0.163) & (0.165) & (0.186) \\ 
  & & & & & & \\ 
 treated & 2.544$^{***}$ & 2.570$^{***}$ & 2.724$^{***}$ & 2.349$^{***}$ & 2.391$^{***}$ & 3.299$^{***}$ \\ 
  & (0.101) & (0.101) & (0.121) & (0.110) & (0.110) & (0.135) \\ 
  & & & & & & \\ 
 sate2 & $-$0.495$^{***}$ & $-$1.488$^{***}$ & $-$0.898$^{***}$ & $-$0.561$^{***}$ & $-$1.497$^{***}$ & $-$1.015$^{***}$ \\ 
  & (0.099) & (0.105) & (0.110) & (0.104) & (0.110) & (0.114) \\ 
  & & & & & & \\ 
 sate3 & $-$0.313$^{**}$ & $-$1.150$^{***}$ & 0.050 & $-$0.375$^{***}$ & $-$1.129$^{***}$ & 0.001 \\ 
  & (0.129) & (0.132) & (0.143) & (0.134) & (0.138) & (0.149) \\ 
  & & & & & & \\ 
 post:treated & 0.098 & 0.056 & $-$1.125$^{***}$ & 0.532$^{***}$ & 0.427$^{***}$ & $-$1.280$^{***}$ \\ 
  & (0.134) & (0.134) & (0.148) & (0.149) & (0.149) & (0.171) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Control & All & All & All & -CHN & -CHN & -CHN \\ 
Month FE & No & Yes & Yes & No & Yes & Yes \\ 
Flag FE & No & No & Yes & No & No & Yes \\ 
Observations & 227,873 & 227,873 & 217,467 & 209,135 & 209,135 & 198,729 \\ 
R$^{2}$ & 0.010 & 0.016 & 0.022 & 0.010 & 0.016 & 0.023 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\clearpage

\hypertarget{other-model-specifications-for-purse-seiners}{%
\subsection{Other model specifications for purse
seiners}\label{other-model-specifications-for-purse-seiners}}

We include a second degree polynomial for years as:

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \omega Y_t + \Omega Y_t^2 + \epsilon_{i,t}
\]

\textbackslash{}begin\{table\}{[}!htbp{]} \centering 
\textbackslash{}caption\{\label{tab:purse}Fishing hours from GFW for
purse\_seiners. Asterisks indicate significance levels. Numbers in
parentheses represent heteroskedastic-robust standard errors.\} \label{}

\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{hours} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 post & 8.394$^{***}$ & 9.334$^{***}$ & 14.934$^{***}$ & 17.284$^{***}$ \\ 
  & (0.267) & (0.254) & (0.312) & (0.403) \\ 
  & & & & \\ 
 treated & $-$1.069$^{***}$ & $-$0.770$^{***}$ & $-$0.993$^{***}$ & 0.259 \\ 
  & (0.249) & (0.234) & (0.219) & (0.269) \\ 
  & & & & \\ 
 post:treated & $-$0.701$^{**}$ & $-$0.899$^{***}$ & $-$0.530$^{*}$ & $-$0.580$^{**}$ \\ 
  & (0.308) & (0.295) & (0.283) & (0.292) \\ 
  & & & & \\ 
 Constant & 11.738$^{***}$ & 11.281$^{***}$ & 7.574$^{***}$ & 10.077$^{***}$ \\ 
  & (0.220) & (0.292) & (0.318) & (0.422) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Month FE & No & Yes & Yes & Yes \\ 
Year FE & No & No & Yes & Yes \\ 
Flag FE & No & No & No & Yes \\ 
Observations & 37,840 & 37,840 & 37,840 & 37,840 \\ 
R$^{2}$ & 0.087 & 0.136 & 0.179 & 0.192 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular}

\textbackslash{}end\{table\}

\clearpage

\hypertarget{quarterly-did-interactions}{%
\subsubsection{Quarterly DID
interactions}\label{quarterly-did-interactions}}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-33-1.pdf}
\caption{\label{fig:unnamed-chunk-33}\label{fig:q1}Interaction of quarters
and treatment. Control group is all vessels.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-34-1.pdf}
\caption{\label{fig:unnamed-chunk-34}\label{fig:q2}Interaction of quarters
and treatment. Control group is vessels from PNA.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-35-1.pdf}
\caption{\label{fig:unnamed-chunk-35}\label{fig:q3}Interaction of quarters
and treatment. Control group excludes Chinese vessels.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-36-1.pdf}
\caption{\label{fig:unnamed-chunk-36}\label{fig:q4}Interaction of quarters
and treatment. Control group is japanese vessels.}
\end{figure}

\hypertarget{year-month-did-interactions}{%
\subsubsection{Year-month DID
interactions}\label{year-month-did-interactions}}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-37-1.pdf}
\caption{\label{fig:unnamed-chunk-37}\label{fig:ym1}Interaction of
year-month and treatment. Control group is all vessels.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-38-1.pdf}
\caption{\label{fig:unnamed-chunk-38}\label{fig:ym2}Interaction of
year-month and treatment. Control group is vessels from PNA.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-39-1.pdf}
\caption{\label{fig:unnamed-chunk-39}\label{fig:ym3}Interaction of
year-month and treatment. Control group excludes Chinese vessels.}
\end{figure}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-40-1.pdf}
\caption{\label{fig:unnamed-chunk-40}\label{fig:ym4}Interaction of
year-month and treatment. Control group is japanese vessels.}
\end{figure}

\hypertarget{effort-redistribution}{%
\subsection{Effort redistribution}\label{effort-redistribution}}

\hypertarget{rasterized-regions}{%
\subsection{Rasterized regions}\label{rasterized-regions}}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-41-1.pdf}
\caption{\label{fig:unnamed-chunk-41}\label{fig:raster_rgn}Rasterized
regions. Each color (number) indicates a distinct region.}
\end{figure}

\hypertarget{fishing-raster}{%
\subsubsection{Fishing raster}\label{fishing-raster}}

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-42-1.pdf}
\caption{\label{fig:unnamed-chunk-42}\label{fig:fishing_raster_full}Yearly
rasters of fishing hours for each group.}
\end{figure}

\clearpage

\end{document}