### 2019-01-27 docs/revenue_loss.tex
Added new missing tables from John

\begin{table}[htbp] \centering 
	\caption{Estimated revenue losses under different scenarios of PNMS (in USD)} 
	\label{tab:revenue_loss}

\begin{tabular}{l*{5}c}

\hline
\hline
Scenario	&	PS VDS	&	LL VDS	&	Export tax	&	Total revenue loss	\\
\hline
	1	&	0	&	-2,100,000	&	-482,236	&	-2,582,236	\\
	2	&	-3,150,000	&	-2,100,000	&	-482,236	&	-5,732,236	\\
	3	&	-5,600,000	&	-2,100,000	&	-482,236	&	-8,182,236	\\
	4	&	-8,750,000	&	-2,100,000	&	-482,236	&	-11,332,236 	\\

	\hline 
	\hline

\end{tabular}

\end{table}