### 2019-03-01 docs/PIPA_Palau022819.tex
Updated fig 5 to include reported vs observed

\documentclass[9p,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\usepackage{placeins}

\title{Large-Scale Marine Protected Areas in the World's Largest Tuna Fishery}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Juan CarlosVillase\~{n}or-Derbez}
\author[b]{John Lynham}

\affil[a]{Bren School of Environmental Science \& Management, University of California at Santa Barbara, Santa Barbara, CA}
\affil[b]{Department of Economics, University of Hawaii at Manoa, Honolulu, HI}

% Please give the surname of the lead author for the running footer
\leadauthor{Villase\~{n}or-Derbez}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{The oceans are becoming increasingly crowded, with different activities competing for space, but little research has been done to understand how activities displace each other. Marine Protected Areas are inherently spatial and have the sole objective of conserving bounded waters through the displacement of fishing effort. Our work shows that displacement of fishing effort occurs, but that it comes at a cost to countries that would have otherwise been able to charge a price for vessels to fish in the now-protected areas. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020 and the expansion of offshore aquaculture, it is important that the spatial displacement of economic activities is fully considered.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally to this work}
\authordeclaration{The authors declare that they have no conflicts of interest}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: juancarlos@ucsb.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Marine Spatial Planning $|$ Fisheries $|$ Marine Conservation}

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. The fishery in the region operates under a Vessel-Day Scheme (VDS), which sells access rights that allow vessels to fish in PNA waters. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: The Phoenix Islands Protected Area (PIPA, 397,447 km\textsuperscript{2}), effectively excluding all tuna purse seining activities. Such an intervention is likely to have effects on vessel spatial distribution and behavior, as well as induce costs (due to the reduction in fishable area) to Kiribati. We use identification of fishing activity via Automatic Identification Systems and causal inference techniques to evaluate the effect of PIPA on vessel distribution, behavior, and costs to Kiribati and the PNA. We find a crowding effect within PNA waters after the implementation of the protected area. Vessels continue to fish with similar intensity after the implementation. In the first year of the closure, there is no drop in total fishing effort within Kiribati's EEZ and a reported increase in revenue from access rights sold. However, from 2016 onwards there is a noticeable drop in fishing effort within Kiribati and a reported drop in VDS revenue. At the same time, fishing effort increases in other parts of the PNA. This redistribution of fishing effort eventually results in a reduction of 2,310 vessel-days in Kiribati, which represents a loss of \$27.7 million USD; similar estimates are obtained when looking at country-level license fees revenues directly (\$30.5 million USD). We use our results to inform predictions of the impacts of a proposed Large-Scale Marine Protected Area (LSMPA) in Palau (another PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. PNA members who indirectly benefit from MPAs should consider mechanisms that reward such conservation actions.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}\label{introduction}

\dropcap{H}umans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and establish causal links between past management interventions and their outcomes \citep{burgess_2018}. One of the most notable spatial interventions is the creation of no-take Marine Protected Areas (MPAs), which seek to conserve the environment by eliminating fishing effort within their waters.

Global international goals aim to protect 10\% of the ocean environment by 2020. In an effort to meet this target, there has been a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large-Scale Marine Protected Areas (LSMPAs), some of which occur in the world's largest fishing grounds for tuna. While today a small number of LSMPAs make up at least 80\% of the managed areas in the ocean \citep{toonen_2013}, very little is known about their human dimensions and implications for fisheries \citep{gray_2017}. Furthermore, most of the research on LSMPAs has focused on their potential ecological benefits, but have left aside the economic implications. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure and represent large opportunity costs \citep{smith_2003,smith_2010}.

The Phoenix Islands Protected Area (PIPA) in Kiribati is one of the most notable Large-Scale Marine Protected Areas. Implemented on January 1st of 2015, PIPA closed an area of 397,447 km\textsuperscript{2} to fishing and was implemented within an area that produces approximately 50\% of the world's tuna. Tuna purse seine fisheries in the region are collectively managed under a Vessel-Day Scheme (VDS) by nine countries commonly refered to as the Parties to the Nauru Agreement (PNA). Members include the Federated States of Micronesia, Kiribati, the Marshall Islands, Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu; Tokelau joined the PNA group in 2012 and started selling access rights in 2013 (Figure \ref{fig:PNA_map}). The Nauru Agreement regulates access of foreign vessels (\emph{i.e.} those from non-PNA countries). Holding 80\% (14.6 million km\textsuperscript{2}) of historical skipjack tuna purse seine grounds within their Exclusive Economic Zones (EEZ), PNA countries have achieved greater bargaining power when providing fishing access to foreign fleets \cite{havice_2010}. The revenue from access fees may represent up to 50\% of government revenue for some of the members.

A spatial closure of such dimensions is likely to cause changes in spatial distribution and behavior of fishing vessels. For example, the anticipation of LSMPAs may lead to preemptive overfishing, which will likely erodes or delay the expected benefits of an intervention \cite{mcdermott_2018,hanich2018unraveling}. After a spatial closure, users can exhibit idiosyncratic responses driven by different incentives \cite{cabral_2017}. Moreover, this reduction in total fishing area within one country's EEZ is likely to result in selling less fishing licenses. While no studies have assessed the implications of PIPA, other PNA members have pledged the implementation of LSMPAs by 2020 (i.e. Palau). We evaluate the behavioral responses and spatial redistribution of the industrial tuna purse seine fleet resulting from the implementation of the Phoenix Islands Protected Area, and quantify its economic ramifications and impacts to Kiribati. We use the same data to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are controled by PNA countries, where the largest tuna fisheries occur.

We use identification of fishing activity via Automatic Identification Systems (AIS) to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation. The remaining 28 vessels never fished in PIPA waters. We refer to these groups as the treated and control groups. The group with the remaining 221 vessels contains vessels that were not observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

\begin{figure}
\centering
\includegraphics{img/LSMPAs_map.pdf}
\caption{\label{fig:LSMPAs_map}Large Scale Marine Protected Areas. The map shows all areas larger than 30,000 Km\textsuperscript{2}. The Phoenix Islands Protected Area is shown in red.}
\end{figure}

\section{Results}\label{results}

\subsection{Crowding effect}

We first inspect the crowding effects that may arise due to the net reduction in fishing area. We produced 1-degree rasters of monthly fishing effort for our treatment and control groups, and calculated two indices of spatial overlap between them: 1) the number of cells that had fishing activity from treated and control vessels for each month and 2) the correlation of presence/absence of fishing events between both groups over one month. We find that the two fleets significantly interact more with each other after the implementation of PIPA (Table \ref{tab:sp_corr} Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and spatial correlations increase by a factor of four and three, respectively. This increase in crowding is likely to increase the encounter rates with other vessels, and reduce the efficiency of fishing operations. This might cause vessels to leave their current fishing grounds and re-optimize their spatial effort, leading to a subsequent decrease as the crowding measures return towards pre-implementation levels.

\begin{figure}[H]
\centering
\includegraphics{img/sp_corr.pdf}
\caption{\label{fig:sp_corr}Number of cells that had treated and control vessels (A) and spatial correlation in the presence-absence of treated and control vessels per cell (B). The solid lines represent the 4\textsuperscript{th} degree polynomial fit reported in \ref{tab:sp_corr}}
\end{figure}

\subsection{Behavioral changes}

The behavioral responses that vessels can have to a spatial closure may occurr in different ways. For example, displacement to new fishing grounds may represent a cost as fishers search the ocean to identify the most suitable fishing spots. This may result in increased fuel and labor costs. For every vessel in each group, we calculate eight key measures that could capture responses to spatial closures: daily fishing hours, daily non-fishing at-sea hours, the proportion of fishing to non-fishing hours at sea, daily distance traveled, daily mean distance from shore of fishing events (km), daily mean distance from port of fishing events (km), as well as monthly hours spent in PNA waters and Kiribati waters (Fig. \ref{fig:all_panels}). We leverage our Before-After-Control-Impact (BACI) design and implement a log-linear difference-in-differences analysis to evaluate how these measures change for treated vessels after implementation of PIPA, relative to the trends observed for control vessels (see the Methods section for our empirical specification).

We find no evidence of treated vessels fishing for more hours after PIPA implementation, and a slight decrease of fishing hours relative to total at-sea hours ($p < 0.01$; Table \ref{tab:main_DID}). Treated vessels traveled less, with fishing events occurring closer to shore and closer to port following PIPA implementation. These changes in distance from shore and port are likely caused by redistribution, as we observe that treated vessels fish 56.5\% and 39.9\% less in Kiribati and PNA waters, compared to the trend observed for control vessels ($p < 0.01$). We do not observe a statistically significant increase in fishing hours on the high seas.

The fact that the amount of time that vessels spent at sea (fishing and non-fishing) did not decrease, and that their time spent in Kiribati and PNA waters decreases suggests that these vessels have re-distributed elsewhere. This is consistent with the eventual decrease in crowding meassures.

\subsection{Economic impacts}

The crowding effect combined with the reduction of hours spent in Kiribati and PNA waters overall suggests that treated vessels have redistributed elsewhere, meaning that they no longer buy access fees from PNA countries. To quantify the potential impacts of this leakage, we estimate the total annual vessel-days received by all PNA countries by each group of vessels (Fig. \ref{fig:all_PS_VDS_year}). In this case we look at all 313 vessels, but continue to group them as treated (n = 64) and control (n = 28) vessels, as well as other vessels (n = 221). We find that the treated vessels spent 2,310 less days in PNA waters when comparing their values for 2015 and 2016, while control and other vessels increased their activities in PNA waters by 5,489 and 4,447 days for the same period, respectively (Fig. \ref{fig:included_PS_VDS_year_DiD}). The 2,310 vessel-day reduction by treated vessels corresponds to \$27.7 million USD at an average price of \$12,000 USD per vessel-day. Looking at the total anual vessel-days allocated by all vessels to all countries, we see that the largest reductions ocurr for Kiribati, while Papua New Guinea exhibits a proportional increase (Fig. \ref{fig:all_PS_VDS_cty_year}). Furthermore, in 2017 treated vessels exhibit the lowest vessel-day allcoation to PNA waters at just 10,026 vessel-days. This is 2,526 vessel-days lower (\~\$30.3 million USD) than pre-implementation levels of 2012 - 2013 (mean $\pm$ sd of 12,552.46 $\pm$ 1,396.46; we exclude 2014 due to the blue paradox effect \cite{mcdermott_2018}). % Do you think we need to come up with an identification strategy for this?. I think a simple DID doesn't work, because the post-trend would assume that there is no upper bound to effort, right? Current DiD estimate of these data suggests a 8,700 vessel-days reduction; or 7,400 vessel-days if we account for the blue paradox with some fixed effects. These values would represent forgone profits of 104 to 88 million USD. Again, this is if the treated vessels were allowed to maintain an upward sloping trend as observed with the control. I am also not sure if this trend is true, or just an artifact of GFW. Could we de-trend it first so that both lines are flat before, and only treated vessels show a decrease? In that setting we can then do a DiD and calculate the gap more accurately.

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_year.pdf}
	\caption{\label{fig:all_PS_VDS_year}Observed vessel-days for all countries by treated, control, and excluded vessels.}
\end{figure}

\begin{figure}[H]
\centering
	\includegraphics{img/included_PS_VDS_year_DiD.pdf}
	\caption{\label{fig:included_PS_VDS_year_DiD}Vessel days spent inside A) PNA waters and B) Kiribati waters by group of vessels.}
\end{figure}

We compliment our analysis of change in observed vessel-days by looking at country-level data. Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA\footnote{https://www.ffa.int/node/2050}) where annual revenues from license fees are reported for each country (2008 - 2016; Fig. \ref{fig:revenues}A). We find that by 2016, Kiribati's revenue went from \$148.8 million USD in 2015 to \$118.3 million USD in 2016, representing a decrease of \$30.5 million USD. We then calculate total annual revenues to all PNA members (Fig. \ref{fig:revenues}B). To estimate what would have been the total revenue to PNA members, we use a local polynomial regression fit to obtain a trend line for the total revenues between 2008 and 2015, and extrapolate the trend to obtain an expected value for 2016. Our model suggests that total revenues to all PNA members in 2016 should be in the order of \$463.5 million USD, but the FFA data show only \$423.5 million, suggesting that potential losses to the entire PNA are in the order of \$40 million USD.

\begin{figure}
\centering
	\includegraphics{img/revenues.pdf}
	\caption{\label{fig:revenues}Financial indicators for PNA countries. A) Revenue from fishing licenses fees, B) Total annual revenues for all 9 PNA countries with dashed line showing the local polynomial regression fit to the 2008 - 2015 data and the red dot representing the prediction from this regression. Vertical dashed line in both plots denotes implementation of PIPA.}
\end{figure}

Catches for each country's EEZ for the 1997 - 2016 period were also obtained from the FFA (Fig. \ref{fig:catches}). Catches in Kiribati waters decreased from 24,051 to 12,894 tonnes between 2015 and 2016 (46.3\% decrease). Similar decreases were observed for The Federal States of Micronesia (60.9\%), Papua New Guinea (43.4\%) and the Solomon Islands (58.5\%). In contrast, Tokelau (due south of Kiribati) showed a 22.3\% increase in catches over the same period.

\subsection{Potential Revenue Loss for Palau}

On October 28, 2015, the President of Palau, H.E. Tommy E. Remengesau Jr., signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 km\textsuperscript{2} to commercial fishing activities, creating the 14th largest protected area in the world. The sanctuary will fully protect about 80 percent of the nation’s EEZ. To prepare for full enactment of the PNMS, the act stipulates a “winding down” period, in which baseline vessel days (i.e., the number of vessel days used in 2014) were reduced by 20\% in 2016; and an additional 10\% from baseline in each subsequent year until full enactment in 2020. This appears to be occurring for the longline VDS, but it is uncertain whether this is being followed for the purse seine VDS, because purse seine VDS can be transferred for use in other EEZs. There has been no official statement on what will happen to Palau’s VDS upon full implementation of the PNMS. However, there is a sense that Palau will be able to keep its allotment come 2020 (Hanich, pers. comm.). In the 2015 Micronesian Presidents Summit, a letter was drafted by heads of state, calling on PNA members to be supportive of Palau as they moved forward with the PNMS Act \cite{Senase2015}. Further, other PNA members have not been penalized for other protected area closures (e.g., PIPA) (Hanich, pers. comm.).

Table \ref{tab:revenue_loss} presents estimates of the potential revenue losses following full enactment of the PNMS under four different scenarios. In Scenario 1, Palau is able to keep its current allotment of purse seine vessel days (700) and is able to sell them for a similar price to what it is currently selling them to the US for (\$12,500/day). In Scenario 2, Palau is able to keep its current allotment of purse seine vessel days (700) to transfer to other parties at the current benchmark price (\$8,000/day). Scenario 2 is likely if Palau retains its current PAE, but the US no longer purchases days.  It should be noted that if PAE continues to be calculated based on effort and biomass, and if Palau continues to be allocated vessel days, its PAE will decrease as effort in its EEZ reaches zero. In Scenario 3 and 4, Palau loses all of its PS vessel days, at \$8,000/day and \$12,500/day, respectively. In all scenarios, all longline vessel day and export tax revenues are lost. Longline vessel day loss is calculated using an average value of \$200 for 10,500 days. Export tax loss is calculated given the average tax revenue from 2012-2014 (\$482,236  from \cite{Gillett2016}).

\input{revenue_loss}

\section{Discussion}

Our findings provide insights into the effect that LSMPAs can have on vessel behavior and the redistribution of fishing effort. We find a short-term crowding effect after the implementation of the protected area. Behavioral analysis shows that the implementation of PIPA had little effect on the total fishing effort exerted by purse seiners. This, combined with our analysis of crowding effects that eventually return to pre-implementation levels, suggests that vessels redistribute elsewhere. Our analysis suggests that vessels that fish elsewhere represent a loss in profits of around \$27.7 million USD to Kiribati. A thought experiment for Palau's Marine Sanctuary suggests losses in profits of up to \$11 million USD. Here, we discuss the implications of our findings and possible shortcommings in our analyses.

Previous studies on protected areas around Pacific islands suggest that vessels move to distant places, which might be translated as increased costs \citep{stevenson_2013}. Others have used similar satellite-tracking systems to show that fishing effort accumulates near the edges of spatial closures, yielding greater catches over time \citep{murawski_2005}. But these vessel tracks do not cover the pre-reserve period, making it difficult to identify the contribution of spatial closures to the observed spatial distribution of fishing vessels. Recent work by \cite{elahi_2018} identified that total fishing effort in a focal region where a short-term MPA was implemented showed little change, likely indicating that fishers redistributed fishing effort to compensate for the reduction in available space. Our data, which is assembled in a similar way, allows us to make similar inferences about the unobserved change in aggregate fishing effort and its spatial redistribution.

A major shortcoming of our analysis is that we do not observe catches or revenues at he vessel level, which ultimately are the factors that guide the decision making process of profit-maximizing agents. Therefore, it is difficult to know whether the small change in fishing hours and redistribution represents a positive or negative impact. An additional factor that we are yet to test is the change in non-fishing hours spent at sea. It is plausible that fishing hours remain constant, but vessels have to increase search time. Upon being relocated, fishers may not identify the best fishing grounds as easily as before, and therefore invest a greater proportion of their time searching for their catch. Further analysis of temporal trends in non-fishing hours, as well as distance traveled should provide us with insights as to why fishers reduced fishing hours. However, the widespread footprint of the treated fleet may suggest that they are well acquainted with the region (see Figure \ref{fig:fishing_raster}), and it's unlikely that they would need to invest much time to identify new fishing grounds.

A growing body of literature suggests that closing the high seas to all fishing could increase fishery yields and profitability of fisheries, with negligible costs to food security \citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}. Our work suggests that the implementation of LSMPAs has little impact on total fishing effort, but that it may result in large losses to countries. We also show that spatial closures lead to short-term crowding effects, which causes vessels to redistribute to areas close by. Such management interventions should consider how fishing effort will change in space and through time, and the ecological implications of this redistribution to ensure that fishing effort is not just displaced elsewhere. While LSMPAs can provide a wide range of benefits, their implementation must be accompained with traditional fisheries management to maximize effectiveness, consider the opportunity costs of such closure, and identify sustainable mechanisms that would compensate losses and incentivize marine conservation.

\section{Methods}

\subsection{Data}

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. GFW then uses machine learning algorithms (convolutional neural networks) on the broadcast messages to infer what type and location of fishing events \citep{kroodsma_2018}.

% Numbers in this paragraph to be updated
% Our analysis focuses on tuna purse seine vessels, the most important fishery for PNA countries. We identify a total of 186 purse seiners that fished in PNA waters at least once before 2015. Vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses, leaving us with 125 vessels. We identify 90 vessels that have fished inside PIPA at least once since 2012; 72 did so before the announcement of the closure (\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}). From these, we only observe before and after data for 64 tuna purse seiners.
%On the other hand, 45 vessels never fished inside PIPA during the 2012 - 2015 period, but we only observe befor and after data for 28 of these.

Our treatment group contains all purse seiners (n = 64) that fished within PIPA at least once before the announcement, and that continued to fish elsewhere after the January 2015 implementation. Vessels in the control group meet the following two conditions: i) never fished within PIPA waters from 2012-2015, and ii) vessels have fished in surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA closure (n = 28). Together, these vessels represent more than 20 million georeferenced positions for which we know activity (fishing or not fishing). We include three additional control groups as a robustness check. The first group excludes all Chinese vessels, the second group excludes all PNA vessels, and the third group excludes US and Taiwanese vessels. Our main definition of treatment and control groups leaves us with 64 treated and 28 control vessels, which have just over 36 million observations.

\subsection{Analyses}

\subsubsection{Crowding effect}

If the implementation of the reserve induces a crowding effect, we would expect to observe no trend before the implementation of the reserve. The implementation would lead to an increase in these measures, which,  if vessels avoid crowding, should then come back down. We therefore expect to have three inflection points: 1) in the transition between no trend and initial crowding due to MPA implementation, 2) When the crowding has reached its maximum and starts to decrease, and 3) when the decrease levels off, pressumably to pre-MPA levels. For this reason, we fit a 4th degree polynomial to our monthly indices. We do so by centering our time series of crowding indices on the day of implementation. Our explanatory variable is therefore the number of months ($M$) before or after the implementation. For example, since PIPA was implemented in January 1st of 2015, December of 2014 has a value of -1, and Feb of 2015 would receive a value of 1.

\begin{figure}
\begin{align}
y_t = \alpha + \beta_1 M_t + \beta_2 M_t^2 + \beta_3 M_t^3 + \beta_4 M_t ^4 + \epsilon_t
\label{eqn:sp_corr}
\end{align}
\end{figure}

\subsubsection{Behavioral changes}

We then focus on identifying the response of vessels to the PIPA closure. We use daily fishing and non-fishing hours, daily proportion of fishing vs. non-fishing hours, daily distance traveled (km), distance from shore (km) and distance from home port(km) for fishing and non-fishing events, and proportion of total fishing hours allocated to Kiribati waters and PNA waters as our main outcomes of interest. We compare these outcomes before and after the implementation of PIPA using a Difference-in-Differences approach. Our main specification is the following:

\begin{figure}[H]
\begin{align}
log(y)_{i,t} = \alpha + \beta_1 P_t + \beta_2 T_i + \beta_3 P_t \times T_i + \phi_t + \gamma_i + \epsilon_{i,t}
\label{eqn:did}
\end{align}
\end{figure}

Where $log(y_{i,t})$ is the log-transformed outcome of interest for vessel $i$ on day $t$. A dummy variable $Post_t$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 for all dates following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the treatment ($Treat_i = 1$) or control ($Treat_i = 0$) group. $\alpha$ is the standard intercept term, $beta_1$ captures the temporal trend, $beta_2$ captures the initial difference between treated and control groups, and $beta_3$ is our parameter of interest: the Difference-in-Differences estimate capturing the treatment effect. Finally, $phi_t$ and $\gamma_i$ represent month and flag dummies that account for seasonality or country-level management interventions. We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results\footnote{I actually need more time to run these, but I don't think they'll change}.

All regression coefficients were estimated via ordinary least squares, and heteroskedasticity-robust standard errors were calculated. All analyses were performed in R version 3.5.1 \citep{rcore_2018}. Raw data and code used in this work are available on \href{https://github.com/jcvdav/MPA_displacement}{github}.

\subsubsection{Revenues}

Information about FFA reports here

\section{References}

\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%		 					 %%%%%%%%%%%%%
%%%%%%%%%%%%%		 BEGIN APPENDIX		 %%%%%%%%%%%%%
%%%%%%%%%%%%%		 					 %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\onecolumn

\FloatBarrier

\section{Appendix}\label{appendix}

\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\section{Supplementary tables and figures}

\input{tab/main_DID}

\clearpage

\input{tab/sp_corr}

\clearpage

\begin{figure}
\centering
\includegraphics{img/PNA_map.png}
\caption{\label{fig:PNA_map}Map of the Exclusive Economic Zones (EEZs) of the region of interest. Countries that belong to the PNA are shown in blue, while empty polygons indicates all others. A red line indicates the Kiribati EEZ, and a solid red polygon delineates PIPA.}
\end{figure}

\clearpage

\begin{figure}
\centering
\includegraphics{img/all_panels.pdf}
\caption{\label{fig:all_panels}Time series showing monthly averages for eight variables of interest: A) Fishing hours, B) Non-fishing hours at-sea, C) Proportion of fishing hours to total hours at-sea, D) Distance traveled, E) Mean distance from port for fishing evetns, F) Mean distance from shore for fishing events, I) Monthly hours spent in Kiribati waters, J) Proportion of fishing hours spent in PNA waters.}
\end{figure}

\clearpage

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_cty_year.pdf}
	\caption{\label{fig:all_PS_VDS_cty_year}Observed vessel days by country and year}
\end{figure}

\clearpage

\begin{figure}
\centering
	\includegraphics{img/hist_kir_fishing.pdf}
	\caption{\label{fig:hist_kir_fishing}Ridgeplot for the density of the \% of total fishing hours that take place within Kiribati EEZ waters by year for treated vessels where the unit of observation is an individual vessel.}
\end{figure}

\clearpage

\begin{figure}
\centering
	\includegraphics{img/catches.pdf}
	\caption{\label{fig:catches}Financial indicators for PNA countries. A) Annual catches by EEZ and, B) Annual value of catches by EEZ. Vertical dashed line in both plots denotes implementation of PIPA.}
\end{figure}

\clearpage

\begin{figure}
\centering
\includegraphics{img/other_specifications.pdf}
\caption{\label{fig:other_specifications}Alternative difference-in-differences estimates
for our variables of interest using different model specifications. Table \ref{tab:main_DID}
reports estimates for models with month and flag fixed effects (\emph{i.e.} green dots).}
\end{figure}

\clearpage

\begin{figure}
\centering
\includegraphics{img/fishing_raster.png}
\caption{\label{fig:fishing_raster}Yearly spatial distribution of fishing effort by treated and control vessels. Color corresponds to \% of total fishing effort in each panel. Red polygons show LSMPAs in the region.}
\end{figure}

\clearpage

\section{Palau National Marine Sanctuary}\label{PNMS}

\begin{figure}
	\includegraphics[width=0.8\textwidth]{img/Color_V1.png}
	\caption{Map of the Four Protected Areas}
	\label{fig:maps}
\end{figure}

\clearpage

\section{Alternative Controls}
\subsection{Excluding all Chinese vessels}

\input{tab/DID_without_CHN.tex}

\clearpage
\subsection{Excluding all PNA vessels}

\input{tab/DID_without_PNA.tex}

\clearpage

\subsection{Excluding all USA and TWN vessels}

\input{tab/DID_without_USA_TWN}

\clearpage

\section{Additional methods}

This section is divided into two main parts. First, we provide a general description of AIS data and the process of identification of vessel-level fishing events performed by Global Fishing Watch.\footnote{Global Fishing Watch: \url{globalfishingwatch.org}} Alongside, we describe the subset of data used in our analysis. We also point out possible shortcomings in the data, or factors that must be considered in the analysis. We then move on to explain our empirical strategy for the identification of behavioral changes and the redistribution of fishing effort.

\subsection{Data}

The amount of data gathered by GFW is dependent on the number of antennas and satellites that can receive signals. The total satellite count increased from 3 to 6 on June 1\textsuperscript{st} 2014 , and then from 6 to 10 on January 1\textsuperscript{st} 2016. This causes an increase in the number of \emph{received} AIS messages (\emph{i.e.} points), and therefore an apparent increase in the number of vessels. The addition of new satellites affects all vessels in the same way.

\input{tab/baci_n_s}

%\section{Old Text}

%\subsection{Large-Scale Marine Protected Areas}\label{lsmpas}
%
%The cutoff at which an MPA is considered to be a LSMPA ranges from areas
%larger than 30,000 km\textsuperscript{2} as defined by
%\cite{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
%as defined by \cite{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
%LSMPAs that meet the latter condition, and are also fully no-take. LSMPAs are often
%implemented in the pelagic environment, where the dominant human
%activity is industrial fishing \cite{gray_2017,kroodsma_2018}. The
%early literature on LSMPAs focused on the inherent challenges and
%difficulties that come with a pelagic environment. \cite{kaplan_2010}
%claimed that very large MPAs would result in excessive opportunity costs
%and that these would be difficult to enforce. \cite{game_2009}
%suggested that most of the challenges could be overcome with the
%incorporation of technology, in what then became known as Dynamic Ocean
%Management \citep{maxwell_2015}.\footnote{See \cite{singleton_2014}, who provide an objective discussion of the pros and cons of LSMPAs.} 
%
%Spatial closures of this magnitude are likely to induce changes in
%fishers' behavior. Theoretical models of fishing effort redistribution
%range from the simplistic assumption that effort inside the bounded
%region disappears, to spatially explicit models that reallocate fishing
%effort based on habitat characteristics, presence of other vessels, and
%expected returns \citep{smith_2003,hilborn_2006}. However, these focus
%on the long term optimal equilibrium, and redistribution of fishing
%effort may not always be optimally distributed, especially over the first few years
%\citep{stevenson_2013}.
%
%The empirical research that has been done in smaller sized MPAs
%suggests that resource users may show idiosyncratic responses. For
%example, \cite{stevenson_2013} show that a network of MPAs displaced
%fishing effort farther away from ports, resulting in higher
%\emph{perceived} costs, and increases in catch per unit effort.
%\cite{cabral_2017} analyze the redistribution of fishing and
%non-fishing vessels following the implementation of a network of MPAs in
%California, and find that dive boats follow a
%fishing-the-line pattern, while some fishing boats follow an ideal free
%distribution. More recently \cite{elahi_2018} used satellite tracking
%data to show that a temporal spatial closure caused trawlers to maintain
%effort but apply it more intensively elsewhere, particularly along the
%borders and closer to shore. The way in which fishers react to a spatial
%closure can have major implications for its outcome
%\citep{smith_2003,hilborn_2006}, highlighting the need to understand how
%fishers react to the implementation of LSMPAs, how fishing effort
%changes, and how it is spatially redistributed. All these studies evaluate relatively small closures within Exclusive Economic Zones, where other regulations exist.
%This may not always be the case for LSMPAs, where often the entire EEZ
%is converted into a LSMPA, leaving fishers with the option of moving to
%the high seas or other countries' EEZs, with potentially very different fishing regulations.
%
%\subsection{Nauru agreement and the Phoenix Islands Protected Area}
%
%The cooperation that emerged under the Nauru Agreement allowed for subsequent
%agreements that strengthened fisheries management, like the Palau
%Agreement, which limited the number of purse seiners at 205 vessels from
%1995-2007.\footnote{See \cite{havice_2010} for a detailed description of the Nauru, Palau, and Federal States of Micronesia agreements.} However, the most notable regulation is
%the approach to manage fishing effort: a Vessel Day Scheme (VDS)
%implemented in 2007 \citep{havice_2013}. This effectively modified how
%fishing effort was managed, from total number of vessels (under the Palau
%Agreement) to total vessel-days. 
%Under the purse seine VDS, a total number of annual vessel days for the entire fishery is agreed upon by the PNA parties. Each party’s allowable effort (PAE) is then calculated based on historic effort and biomass within each party’s EEZ. Sixty percent of the PAE is calculated based on EEZ effort over the last seven years and 40\% of the PAE is calculated based on the 10-year average of each country’s share of estimated skipjack and yellowfin biomass within its EEZ.\footnote{This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}}. A minimum benchmark fee is set for purse seine vessel days which each party can transfer (\emph{i.e.} sell to another PNA member) or sell to the highest bidder (\emph{i.e.} sell to a fishing company). Vessel days can be transferred to fish in any PNA party’s EEZ, without penalty to the transferred parties' allocations \cite{PNA2016}.\footnote{The process of transferring days between parties is described in Article 7 of the Management Scheme \cite{PNA2016}.} The mean value of a vessel day has steadily increased since 2007 \citep{havice_2013}. 
%Although detailed records on sales or transfers of vessel days are not publicly available \citep{havice_2013,yeeting2018stabilising}, a minimum benchmark fee of \$8,000/day was set from 2015 \citep{PNA2014a}.
%
%\cite{yeeting2018stabilising} summarize the PNA implementing arrangements as follows: “foreign vessels [are required] to be registered and licensed, report catches, maintain log books, allow observers on board and maintain transparency over their fishing activities.” Further, vessels registered with the Vessel Day Scheme are required to have Automatic Location Communicators (ALC) or Mobile Transceiver Units (MTU) that transmit their locations at least once per hour while they are within the VDS Management Area \citep{PNA2016}. Every entire day that a vessel is within the VDS Management Area is counted as a vessel day used, unless the vessel reports a “no fishing day” (e.g., travel, maintenance, etc.) and periods of less than 24 hours are counted as partial days \citep{PNA2016}.  As described in the Palau Arrangement, fishing days by vessels shorter than 50m (longer than 80m) count as 0.5 (1.5) vessel days. Countries are responsible for ensuring registered vessels comply with the implementing arrangements and that they stay within their PAE. Parties that exceed their PAE should reportedly be penalized by reductions to the following year’s PAE \citep{PNA2016}. There are some criticisms of the PNA PS VDS claiming that there is a lack of monitoring, compliance, and transparency that could hinder the future success of the program \citep{Arnason2014,yeeting2018stabilising}. While the effectiveness of this scheme has been
%debated in terms of meeting fishery management and conservation
%objectives, the licensing significantly contributes to the economy of
%these island nations \citep{havice_2010}.
%
%The main tuna species caught in the region are skipjack
%(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
%albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
%From these, the first two are amongst the top-10 species represented in
%global fisheries production statistics, with 2016 catches increasing
%relative to the 2005-2014 average \citep{fao_2018}. This region of the
%Pacific has historically accounted for a large portion of tuna catches
%\cite{aqorau_1997}. Today, the PNA controls close to 50\% of global
%skipjack tuna production \cite{pna_website_2018} and the combined area of all EEZs involved is 14.6 million km\textsuperscript{2}, larger than the land mass of the United States of America. A large portion of global skipjack
%catch derives from purse seine vessels licensed under the VDS.
%Fishing vessels from Australia, New Zealand, China, France, Korea,
%Japan, the Philippines, Taiwan, and the United States participate in the
%purse-seining VDS.
%
%One of the most notable and recent management interventions in the
%region is the implementation of the Phoenix Islands Protected Area (PIPA)
%by the government of Kiribati. PIPA was first declared in 2006, and
%established in 2008 with only 4\% of the area declared as no-take. On
%January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
%expanded to a total area of 397,447 km\textsuperscript{2}, roughly 1.5
%times the size of Ecuador. Figure \ref{fig:PNA_map} shows a map of the
%PNA countries (excluding Palau) and the Phoenix Islands Protected Area.
%
%The closure of such a large area in one of the most important fishing
%regions in the world provides a unique opportunity to evaluate the
%behavioral responses and redistribution of fishing effort by vessels
%that used to fish there. PIPA has been the focus of previous research
%showing that fishing effort is effectively reduced after implementation
%\citep{mccauley_2016,mcdermott_2018}. To this, we pose two questions:
%How did individual vessels respond to the sudden exclusion of such a big
%area? Where did all the vessels go? And what does this imply for revenues from
%selling VDS for both Kiribati and the PNA as a whole? 
%
%A spatial closure might cause fishers to modify
%their behavior as they adapt to a new state of the world. For example, some may have
%to travel further distances to find new fishing grounds, increasing their fuel
%costs. If fishers had developed experience for fishing in particular
%sites, being excluded might impose a learning cost on them, as they
%identify new fishing grounds. This might result in increased search
%times. However, if the area of knowledge of a vessel is significantly
%larger than that of the spatial closure, they might already know other
%places to fish. In the next sections we describe
%the data and methods used to answer these questions.
%
%\subsection{Palau}
%
%\subsubsection{Palau National Monument}

%On October 28, 2015, the President of Palau, H.E. Tommy E. Remengesau Jr., signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 km\textsuperscript{2} to commercial fishing activities, creating the 14th largest protected area in the world. The sanctuary will fully protect about 80 percent of the nation’s EEZ. The remaining 20\% of Palau’s EEZ (close to the most heavily populated islands: Koror and Babeldaob) will become a Domestic Fishing Zone in which traditional and domestic fishing activities will be allowed to provide fish solely for the domestic market (\emph{i.e.} exports of pelagic fish will be prohibited). Alongside these protections, the Palau National Marine Sanctuary (PNMS) Act will support a domestic pelagic fishery. In this section, we use our preceding analysis on the ex-post impacts of PIPA to attempt to make some \emph{ex ante} predictions about the potential impacts of PNMS. We start by giving some background on the fishing that is currently taking place inside the future PNMS.

%\subsubsection{Background on Commercial Fishing in Palau}

%Foreign tuna fishing in Palau’s EEZ began before WWI. Pole and line was the initial gear used to target skipjack tuna by Japanese vessels. Foreign fishing activities stopped during WWII and started again in the 1960s when Japanese fishers returned and the locally based Van Camp Seafood Company carried out fishing with boats crewed by Okinawans. The Japanese also began purse seine fishing in the 1960s and 1970s and limited longline fishing for yellowfin tuna developed during the same time. During the 1980s, longline fishers began to set their lines deeper, shifting their targets to bigeye tuna. During the 1980s, Korean and Taiwanese vessels also began longline fishing in Palauan waters for export to Japan \citep{chapman2000development}. There are currently three transshipment companies operating in Palau.

%Purse seine vessels target schools of skipjack tuna that are then canned. Compared to other EEZs in the Pacific, especially the PNA EEZs, Palau’s EEZ is not known to be a preferred fishing ground for purse seining (Sisior, pers. comm.). This can be seen in Figure \ref{fig:VDS_country_year}. Currently all purse seine boats fishing in Palau’s EEZ are Japanese and they do not land their catch in Palau (Table \ref{tab:palau_purse}).

\input{palau_purse}

%All longline vessels fishing in Palau’s EEZ are foreign owned (Table \ref{tab:palau_long}). Japanese longline vessels do not land their catch in Palau. Locally-based, foreign-owned longline boats are mostly Taiwanese vessels that are contracted by one of the two main fishing companies (Palau International Traders Incorporated (PITI) and Kuniyoshi Fishing Company (KFC)). Ninety-five percent of the catch from PITI and KFC contracted boats is exported to Japan upon landing in Koror (the largest town in and former capital of Palau). Most of the discards are either sold or donated in Palau. A very small portion of the discards is frozen and transported to Taiwan when vessels return to their home ports.

\input{palau_long}

%\subsubsection{Access fees and taxes}
%
%An agreement between Palau and four Japanese fishing associations\footnote{This is a single agreement between Palau and four associations—1)Federation of Japan Tuna Fisheries Cooperative Associations, 2) National Offshore Tuna Fisheries Association of Japan, 3) Japan Far Seas Purse Seine Fishing Association, 4) Federation of North Pacific District Purse Seine Fisheries Cooperative Association of Japan (Performance Audit Report on Managing Sustainable Fisheries (Tuna) 2013).} covered three methods of fishing (longline, purse seine, and pole and line) and allowed Japan to have up to 290 vessels in Palau waters with no limits on catch. Japan paid 4-5\% of catch returns for this access, but there was no way to validate their catch \citep{Tewid2013}. Between 2010 and 2014, Japan paid Palau between \$196,100- \$867,120 annually in access fees \citep{Gillett2016}. Currently, these access fees have been replaced by the PNA vessel day scheme (described earlier).
% 
%There used to be a number of agreements between the Republic of Palau and foreign owned, local based companies, in particular PITI and KFC, to allow longline fishing \cite{Tewid2013}. Vessels under these agreements were internationally-owned, locally-based vessels that paid for annual licenses based on the size of each vessel. These are the vessels shown in Table \ref{tab:palau_long} (excluding the Japanese vessels). So, for example, there were 38 vessels as part of these agreements in 2016. Between 2010 and 2014, licensing fee revenue from these agreements was between \$219,000-\$284,600 per year \citep{Gillett2016}. These access fees have also been replaced by a new longline-specific version of the VDS (described in more detail below).
%
%There are two multilateral treaties that grant preferential access to Palau waters to US and Federated States of Micronesia (FSM) flagged purse seine boats: the US Multilateral Tuna Treaty and the Federated States of Micronesia Arrangement.  The US Treaty was renegotiated in 2016 with stipulations on minimum vessel day fees. Very little purse seine fishing is conducted by FSM and US vessels in Palau’s EEZ, but Palau still receives a portion of these treaty funds. These agreements have been described by some commentators as foreign aid, with little connection to underlying fish stocks \cite{Gillett2016}. 
%
%The export tax for all commercial tuna and billfish, either fresh or frozen, is \$0.35/kg. Average annual revenue from export taxes was around \$500,000 from 2011-2014 \citep{Gillett2016}. 
%Other than the access fees and tax revenue, Palau receives little additional economic benefits from pelagic fisheries, especially considering that very few jobs are created by the fishery. Between 80 to 90 people are employed by offshore fisheries, and of these only around 20\% are Palauans, earning wages that are only 50\% of average wages in Palau \cite{Gillett2016}.
%
%\subsubsection{Parties to the Nauru Agreement and the Vessel Day Scheme}
%
%In the case of Palau, the VDS is run by the Bureau of Marine Resources, within the Ministry of Natural Resources, Environment and Tourism. Palau’s current purse seine VDS allotment is reportedly around 700 vessel days annually (Sisior, pers. comm., \cite{Pojas2018}), although historically this number has been lower (average of 580 days between 2008-2011 \citep{Tewid2013}). Palau sells most of its vessel days to the US for \$12,500 per day (as negotiated in the US treaty), regardless of whether US vessels actually fish in Palau’s EEZ. The remainder of Palau’s purse seine vessel days are sold for between \$9-10,000 each. When vessel days are purchased to transfer to another EEZ, Palau charges a transfer fee (approximately \$500), for administration costs and future opportunity costs (since vessel day allotments are based partly on historical effort within the EEZ). Though records are confidential, it has been publicly reported that one PNA member party that Palau has transferred vessel days to is Papua New Guinea \citep{Tewid2013} and in 2018, vessel days were sold to a fishing company in the Philippines for the first time \citep{Pojas2018}. In 2018, Palau brought in approximately \$9 million USD in VDS revenue (Sisior pers. comm.). 
%
%The longline VDS (LL VDS) is in its infancy and has not yet been fully implemented at the PNA-scale, although several countries are now implementing it at the country-scale. Palau was the first party to implement the VDS for longline fishing in 2017. In 2014, longline boats fishing in Palau’s EEZ fished 10,500 days (Sisior, pers. comm.). Since 2016, the number of allowed longline VDs has been a fraction of the 2014 days, reducing each year as stipulated in the PNMS Act. Palau sells its LL VD for \$150-\$250 (Sisior pers. comm.). Because the LL VDS is still new and has yet to be fully implemented by all parties, LL VD are not currently transferable between PNA member EEZs. 
%
%\subsubsection{Plans for a domestic pelagic fishery}
%
%In the recent past, a single domestic pole and line vessel supplied Palau with an estimated 100mt of catch \cite{Gillett2016}. The vessel ceased operations in recent years when the captain became ill and passed away, but there is talk that it is gearing up to start operations again under new leadership (Sisior, pers. comm.). Otherwise, there are no domestic vessels dedicated to full-time offshore fishing. A number of fishers occasionally troll offshore for tuna and other pelagics, but there are no official estimates of volumes. 
%Local NGOs have sponsored short trainings on small-scale canning and proper techniques for processing high grade tuna to encourage existing fishers to devote more time to offshore fishing. There are several possible scenarios for a domestic fishery (including those featured in a feasibility report by FFA \citep{Skirtun2017}), but there has yet to be any official action towards a particular scenario.

\end{document}