### 2019-03-07 docs/tab/KIR_sp_corr.tex

\begin{table}[H] \centering 
  \caption{\label{tab:KIR_sp_corr}Coefficient estimates for a fourth-degree polynomial fit to the measures of crowding for Kiribati EEZ only. The first five columns represent different specifications for number of cells with presence of both fleets. Columns 6 - 10 show coefficients for the spatial correlation for presence / absence of displaced and non-displaced vessels. The explanatory variable is the number of months before or after implementation of PIPA. Numbers in parentheses are heteroskedastic-robust standard errors. The last column of each group presents fits with only NINO4 anomaly index as an explanatory variable.} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{0.1pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & \multicolumn{5}{c}{Number of cells} & \multicolumn{5}{c}{Pearson's correlation coefficient} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\ 
\hline \\[-1.8ex] 
 Constant & 30.24$^{***}$ & 32.77$^{***}$ & 23.30$^{***}$ & 26.14$^{***}$ & 16.23$^{***}$ & 0.43$^{***}$ & 0.43$^{***}$ & 0.33$^{***}$ & 0.34$^{***}$ & 0.38$^{***}$ \\ 
  & (2.43) & (4.16) & (4.59) & (5.49) & (2.37) & (0.03) & (0.05) & (0.07) & (0.08) & (0.03) \\ 
  & & & & & & & & & & \\ 
 M & 1.29$^{***}$ & 1.34$^{***}$ & 1.00$^{***}$ & 1.06$^{***}$ &  & 0.01$^{***}$ & 0.01$^{***}$ & 0.01$^{*}$ & 0.01 &  \\ 
  & (0.14) & (0.17) & (0.27) & (0.30) &  & (0.002) & (0.002) & (0.01) & (0.01) &  \\ 
  & & & & & & & & & & \\ 
 M $^2$ & $-$0.03$^{***}$ & $-$0.04$^{***}$ & $-$0.02$^{**}$ & $-$0.03$^{***}$ &  & 0.0000 & 0.0000 & 0.0001 & 0.0001 &  \\ 
  & (0.01) & (0.01) & (0.01) & (0.01) &  & (0.0002) & (0.0002) & (0.0002) & (0.0002) &  \\ 
  & & & & & & & & & & \\ 
 M $^3$ & $-$0.001$^{***}$ & $-$0.001$^{***}$ & $-$0.001$^{***}$ & $-$0.001$^{***}$ &  & $-$0.0000$^{***}$ & $-$0.0000$^{***}$ & $-$0.0000$^{**}$ & $-$0.0000$^{**}$ &  \\ 
  & (0.0001) & (0.0001) & (0.0002) & (0.0002) &  & (0.0000) & (0.0000) & (0.0000) & (0.0000) &  \\ 
  & & & & & & & & & & \\ 
 M $^4$ & 0.0000$^{***}$ & 0.0000$^{***}$ & 0.0000$^{**}$ & 0.0000$^{***}$ &  & $-$0.0000 & $-$0.0000 & $-$0.0000 & $-$0.0000 &  \\ 
  & (0.0000) & (0.0000) & (0.0000) & (0.0000) &  & (0.0000) & (0.0000) & (0.0000) & (0.0000) &  \\ 
  & & & & & & & & & & \\ 
 NINO4 &  & $-$3.10 &  & $-$4.19 & 9.93$^{***}$ &  & $-$0.001 &  & $-$0.02 & 0.11$^{***}$ \\ 
  &  & (3.97) &  & (4.16) & (3.33) &  & (0.05) &  & (0.06) & (0.03) \\ 
  & & & & & & & & & & \\ 
 $\sigma_1$ &  &  & 9.40 & 10.33 &  &  &  & 0.12 & 0.13 &  \\ 
  &  &  & (5.92) & (6.35) &  &  &  & (0.09) & (0.09) &  \\ 
  & & & & & & & & & & \\ 
 $\sigma_2$ &  &  & $-$2.46 & $-$3.84 &  &  &  & 0.02 & 0.01 &  \\ 
  &  &  & (5.70) & (5.77) &  &  &  & (0.10) & (0.11) &  \\ 
  & & & & & & & & & & \\ 
\hline \\[-1.8ex] 
NINO4 & No & Yes & No & Yes & Yes & No & Yes & No & Yes & Yes \\ 
Satellites & No & No & Yes & Yes & No &  &  &  &  &  \\ 
AIC & 654.294 & 655.536 & 655.482 & 656.125 & 724.58 & -65.872 & -63.873 & -63.481 & -61.572 & -35.895 \\ 
Observations & 84 & 84 & 84 & 84 & 84 & 75 & 75 & 75 & 75 & 75 \\ 
R$^{2}$ & 0.63 & 0.63 & 0.64 & 0.65 & 0.08 & 0.44 & 0.44 & 0.45 & 0.45 & 0.09 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{10}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}