### 2019-03-07 docs/tab/sp_corr.tex

\begin{table}[H] \centering 
  \caption{\label{tab:sp_corr}Coefficient estimates for a fourth-degree polynomial fit to the measures of crowding for all PNA waters. The first five columns represent different specifications for number of cells with presence of both fleets. Columns 6 - 10 show coefficients for the spatial correlation for presence / absence of displaced and non-displaced vessels. The explanatory variable is the number of months before or after implementation of PIPA. Numbers in parentheses are heteroskedastic-robust standard errors. The last column of each group presents fits with only NINO4 anomaly index as an explanatory variable.} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{0.1pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & \multicolumn{5}{c}{Number of cells} & \multicolumn{5}{c}{Pearson's correlation coefficient} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\ 
\hline \\[-1.8ex] 
 Constant & 78.04$^{***}$ & 84.64$^{***}$ & 60.96$^{***}$ & 66.73$^{***}$ & 80.62$^{***}$ & 0.41$^{***}$ & 0.42$^{***}$ & 0.37$^{***}$ & 0.38$^{***}$ & 0.38$^{***}$ \\ 
  & (5.44) & (8.99) & (15.60) & (15.99) & (7.96) & (0.01) & (0.03) & (0.06) & (0.07) & (0.02) \\ 
  & & & & & & & & & & \\ 
 M & 3.94$^{***}$ & 4.07$^{***}$ & 3.07$^{***}$ & 3.14$^{***}$ &  & 0.01$^{***}$ & 0.01$^{***}$ & 0.01$^{**}$ & 0.01$^{**}$ &  \\ 
  & (0.30) & (0.35) & (0.97) & (1.04) &  & (0.001) & (0.001) & (0.003) & (0.004) &  \\ 
  & & & & & & & & & & \\ 
 M $^2$ & $-$0.005 & $-$0.02 & 0.01 & $-$0.01 &  & $-$0.0001$^{***}$ & $-$0.0001 & $-$0.0001 & $-$0.0001 &  \\ 
  & (0.02) & (0.03) & (0.03) & (0.03) &  & (0.0000) & (0.0001) & (0.0001) & (0.0001) &  \\ 
  & & & & & & & & & & \\ 
 M $^3$ & $-$0.002$^{***}$ & $-$0.002$^{***}$ & $-$0.002$^{***}$ & $-$0.002$^{***}$ &  & $-$0.0000$^{***}$ & $-$0.0000$^{***}$ & $-$0.0000$^{***}$ & $-$0.0000$^{***}$ &  \\ 
  & (0.0003) & (0.0003) & (0.001) & (0.001) &  & (0.0000) & (0.0000) & (0.0000) & (0.0000) &  \\ 
  & & & & & & & & & & \\ 
 M $^4$ & 0.0000 & 0.0000 & 0.0000 & 0.0000 &  & 0.0000$^{***}$ & 0.0000$^{**}$ & 0.0000 & 0.0000 &  \\ 
  & (0.0000) & (0.0000) & (0.0000) & (0.0000) &  & (0.0000) & (0.0000) & (0.0000) & (0.0000) &  \\ 
  & & & & & & & & & & \\ 
 NINO4 &  & $-$8.09 &  & $-$10.39 & 19.76$^{**}$ &  & $-$0.01 &  & $-$0.01 & 0.07$^{***}$ \\ 
  &  & (8.29) &  & (9.48) & (9.41) &  & (0.03) &  & (0.03) & (0.02) \\ 
  & & & & & & & & & & \\ 
 $\sigma_1$ &  &  & 21.32 & 25.10 &  &  &  & 0.06 & 0.06 &  \\ 
  &  &  & (19.49) & (22.49) &  &  &  & (0.08) & (0.08) &  \\ 
  & & & & & & & & & & \\ 
 $\sigma_2$ &  &  & 5.30 & 3.19 &  &  &  & $-$0.01 & $-$0.02 &  \\ 
  &  &  & (18.87) & (18.67) &  &  &  & (0.04) & (0.03) &  \\ 
  & & & & & & & & & & \\ 
\hline \\[-1.8ex] 
NINO4 & No & Yes & No & Yes & Yes & No & Yes & No & Yes & Yes \\ 
Satellites & No & No & Yes & Yes & No &  &  &  &  &  \\ 
AIC & 796.937 & 797.991 & 799.477 & 799.97 & 919.583 & -187.208 & -185.269 & -184.916 & -183.233 & -97.386 \\ 
Observations & 84 & 84 & 84 & 84 & 84 & 84 & 84 & 84 & 84 & 84 \\ 
R$^{2}$ & 0.79 & 0.79 & 0.79 & 0.80 & 0.03 & 0.70 & 0.70 & 0.71 & 0.71 & 0.07 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{10}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}