### 2019-04-22 docs/Model_with_trading.Rmd
Added new version from John

---
title: "trade"
author: "Villaseñor-Derbez J.C."
date: "22/3/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The vessel-day price for parcels that do not implement a reserve ($j = c(2, 10)$) is given by:

$$
\pi(E_j) = pqX_j - \beta c E_j ^{\beta - 1}
$$

We can solve for $E_j$:

$$
\begin{split}
\pi_j + \beta c E_j ^{\beta - 1} &= pqX_j \\
\beta c E_j ^{\beta - 1} &= pqX_j - \pi_j \\
E_j ^{\beta - 1} &= \frac{pqX_j - \pi_j}{\beta c } \\
E_j &= \left(\frac{pqX_j - \pi_j}{\beta c }\right) ^ { \frac{1}{\beta - 1}}
\end{split}
$$
For the parcel that implements the reserve ($i = 1$) the same equations are given by:

$$
\pi_1(E_1) = pqX_1 (\theta + (1 - \theta)(1 - R))- \beta c E_1 ^{\beta - 1}
$$

For simplicity, we can write $\Omega = (\theta + (1 - \theta)(1 - R))$, and the baove becomes

$$
\pi_1(E_1) = pqX_1\Omega - \beta c E_1 ^{\beta - 1}
$$

We can solve for $E_i$ as in the previous casse:

$$
\begin{split}
\pi_1 + \beta c E_1 ^{\beta - 1} &= pqX_1\Omega \\
\beta c E_1 ^{\beta - 1} &= pqX_1\Omega - \pi_1 \\
E_1 ^{\beta - 1} &= \frac{pqX_1\Omega - \pi_1}{\beta c} \\
E_1 &= \left(\frac{pqX_1\Omega - \pi_1}{\beta c }\right) ^ {\frac{1}{\beta - 1}}
\end{split}
$$

The total vessel-days for cells that belong to PNA countries (*i.e.* i = (1, 9)) are constrained at 45,000, which we term $\bar{E}$. Therefore, total effort can be expressed as:

$$
\bar{E} = \left(\frac{pqX_1(\theta + (1 - \theta)(1 - R)) - \pi_1}{\beta c }\right) ^ {\frac{1}{\beta - 1}} + 
\sum_{j = 2}^9\left(\frac{pqX_j - \pi_j}{\beta c }\right) ^ { \frac{1}{\beta - 1} }
$$

This parameterization shows that the cell-level effort is determined by the combination of stock size ($X_j$) and the price $\pi$. Trading between country will continue until all $\pi_i$ reach an equilibrium price of $\bar{\pi}$.