### 2019-04-23 scripts/model/1_checks.Rmd
Changed script into Rmd

---
title: "Untitled"
author: "Villaseñor-Derbez J.C."
date: "22/4/2019"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---
# Setup

## Rmd options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE)

# Clean the environment
rm(list = ls())
```

## Load packages and functions

```{r}
# Load libraries
library(startR)
library(here)
library(cowplot)
library(tidyverse)

# Load functions
source(here("scripts", "model", "grow.R"))
source(here("scripts", "model", "E_vec_fxn.R"))
source(here("scripts", "model", "pi_of_E.R"))
source(here("scripts", "model", "max_Ei.R"))
source(here("scripts", "model", "solve_pi.R"))
```

# Start checks

## Define parameters

```{r}
# Reserve properties
theta <- 1
R <- 0

# Bio
MSY <- 1875600 # 50th percentile from MSY in table 8 of SA (https://www.wcpfc.int/node/27490)
Bmsy <- 1628000 # 50th percentile from SBmsy in table 8 of SA (https://www.wcpfc.int/node/27490)
K <- 6876526 # 50th percentile from SB_f0 in table 8 of SA (https://www.wcpfc.int/node/27490)
Bc_B0 <- 0.51 # 50th percentile from SBlatest/SB_fo in table 8 of SA (https://www.wcpfc.int/node/27490)
Cnow <- 1679444 # Catches now
Bnow <- K * Bc_B0 # current Biomass (2012 - 2015 average)
r <- 0.57 # From fishbase: Prior r = 0.57, 95% CL = 0.41 - 0.78 (https://www.fishbase.in/summary/107#)

# Economic
beta <- 1.3
p <- 1100 # (1447 + 1467) / 2 #mean between thailand and japan values (Value of WCPFC-CA tuna fisheries 2017 report)
E <- 45000
q <- 12 / (0.1 * Bnow) # 2 * (0.8 * Cnow) / (E * 0.8 * Bnow)
c <- 180 # 340
```

### Calculate historical CPUE

One of the first parameters we need to define is initial biomass. To do this, we will calculate CPUE as a proxy of biomass, and then simply distribute total biomass (from the stock assessment) to create a vector of patch-specific current biomass.

```{r}
# Proportion of effort in the high seas
prop_hs <- 0.1
n_patches <- 10
fvec <- c(rep(((1 - prop_hs) / (n_patches - 1)), (n_patches - 1)), prop_hs)
```

#### Effort data

The first step is to load activity data. This dataset contains activity by vessel and EEZ through time. The following code loads the activity dat and calculates hours and proportional hours by year and eez.

```{r}
# Create vector with VDS country iso3 codes
VDS_countries <- c("FSM", "KIR", "MHL", "NRU", "PLW", "PNG", "SLB", "TUV", "TKL")

# Load historical effort received and summarize
act <- readRDS(here("raw_data", "activity_by_vessel_year_eez.rds")) %>%
  filter(
    eez_iso3 %in% VDS_countries,
    best_vessel_class == "tuna_purse_seines"
  ) %>%
  group_by(year, eez_iso3) %>%
  summarize(days = sum(hours_length) / 24) %>%
  arrange(desc(days)) %>%
  mutate(
    days_prop = days / (sum(days) / (1 - prop_hs)),
    c_days_prop = cumsum(days_prop)
  )
```

This graph shows the data generated above

```{r}
ggplot(data = act, mapping = aes(x = year, y = days)) +
  geom_line() +
  facet_wrap(~eez_iso3, scales = "free_y") +
  labs(x = "year", y = "fishing days") +
  ggtitle("Vessel-days per country")
```

```{r}
ggplot(data = act, mapping = aes(x = year, y = days_prop)) +
  geom_line() +
  facet_wrap(~eez_iso3, scales = "free_y") +
  labs(x = "year", y = "fishing days") +
  ggtitle("Vessel-days per country") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "year", y = "Proportion of vessel-days") +
  ggtitle("Relative effort on each country")
```

#### Catch data

We can now use the FFA catch data for skipjack and estimate yearly catches by EEZ.

```{r}
# Read in the catch data
catches <- read.csv(here("raw_data", "FFA", "ps_skj_catches_spatial_timeseries.csv"), stringsAsFactors = F) %>%
  gather(year, catches, -eez_iso3) %>%
  mutate(year = as.numeric(str_remove(year, "X"))) %>%
  filter(year > 2010) %>%
  mutate(
    PNA = eez_iso3 %in% VDS_countries,
    eez_iso3 = ifelse(PNA, eez_iso3, "HS")
  ) %>%
  group_by(year, eez_iso3) %>%
  summarize(catches = sum(catches))
```

```{r}
ggplot(data = catches, mapping = aes(x = year, y = catches)) +
  geom_line() +
  facet_wrap(~eez_iso3, scales = "free_y", ncol = 3)
```

We can now take catches and divide them by effort to obtain annual CPUE for each EEZ.

```{r}
# Join both data
cpue <- left_join(act, catches, by = c("year", "eez_iso3")) %>%
  drop_na() %>%
  group_by(eez_iso3) %>%
  summarize(
    catches = sum(catches),
    days = sum(days)
  ) %>%
  mutate(cpue = catches / days) %>%
  arrange(desc(cpue)) %>%
  mutate(
    cpue_prop = cpue / (sum(cpue) / (1 - prop_hs)),
    cumsum_cpue_prop = cumsum(cpue_prop),
    eez_iso3 = fct_relevel(eez_iso3, "KIR")
  ) %>%
  arrange(eez_iso3)
```

```{r}
left_join(act, catches, by = c("year", "eez_iso3")) %>%
  drop_na() %>% 
  ggplot(mapping = aes(x = days, y = catches, color = year)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 4) +
  labs(x = "Effort (days)", y = "Catches (tonnes)") +
  ggtitle("Relationship between effort and catches")
```

```{r}
fvec_cpue <- c(cpue$cpue_prop, prop_hs)
names(fvec_cpue) <- c(levels(cpue$eez_iso3), "HS")

# Chose a source for the fvec
bvec <- fvec_cpue # From CPUE data
# fvec <- fvec_h # From effort data
# Initial biomass distribution should be based on CPUE
# Then, biomass distribution vector should be based on area by EEZ (10 % to HS)
Xvec <- bvec * Bnow
names(fvec) <- names(bvec)
```

# Run some checks

```{r}
# Is there a value of E for the given possible ranges of Biomass?
tibble(X = seq(1, Bnow, length.out = 100)) %>%
  rowwise() %>%
  mutate(E = sum(max_Ei(p = p, q = q, Xvec = X, theta = theta, R = R, beta = beta, c = c))) %>%
  ggplot(aes(x = X, y = E)) +
  geom_line() +
  geom_vline(xintercept = Bmsy) +
  geom_hline(yintercept = E)
```

```{r}
# What's the aggregate demand for a given value of pi with Bmsy?
# FIrst we need to establish the maximum chargable price to plot all ranges from 0 to max_pi
max_pi <- max((p * q * Bnow * fvec))

check_E_pi <- tibble(pi = seq(0, max_pi, length.out = 100)) %>%
  rowwise() %>%
  mutate(E = list(
    tibble(
      E = E_vec_fxn(p = p, q = q, Xvec = Xvec, pi = pi, theta = theta, R = R, beta = beta, c = c),
      cty = names(fvec)
    )
  )) %>%
  unnest() %>%
  filter(!cty == "HS")
```

```{r}
ggplot(data = check_E_pi, mapping = aes(x = E, y = pi, color = cty)) +
  geom_line(size = 1) +
  geom_vline(xintercept = E, color = "red")
```

```{r}
check_E_pi %>%
  group_by(pi) %>%
  summarise(E = sum(E)) %>%
  ggplot(aes(x = E, y = pi)) +
  geom_line() +
  geom_vline(xintercept = E, color = "red") +
  geom_vline(xintercept = sum(max_Ei(p = p, q = q, Xvec = Xvec, theta = theta, R = R, beta = beta, c = c)[1:(n_patches - 1)]))
```