### 2019-04-25 docs/slides/NCEAS_presentation.Rmd
Added slides for NCEAS

---
title: "Marine protected areas under effort-control management"
author:
  - Villaseñor-Derbez J.C.$^1$
  - Lynham J.$^2$
  - Costello C.$^{1, 3}$
institute: 
   - $^1$Bren School, UCSB
   - $^2$Department of Economics, UHM
   - $^3$emLAB, UCSB
subtitle: "NCEAS-Future Earth Working Group"
date: "May 6, 2019"
output: beamer_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

## Outline

- Scope of the project

- Approach: model and assumptions

- Data

- Generalizability to Palau

## Background

```{r}

suppressPackageStartupMessages({
  library(startR)
  library(here)
  library(sf)
  library(tidyverse)
  library(ggdark)
})

source(here("scripts", "functions", "st_rotate.R"))

rast <- readRDS(here("data", "rasterized_effort_by_region.rds")) %>% 
  mutate(post = fct_relevel(
    ifelse(year < 2015, "Before", "After"),
    "Before")) %>% 
  group_by(post, x, y) %>%
  summarize(days = log10(sum(hours) / 24)) %>%
  ungroup() %>% 
  filter(between(x, 160, 210),
         between(y, -15, 10))

pipa <- read_sf(dsn = here("data", "spatial", "PIPA"),
                layer = "PIPA") %>% 
  st_rotate()

pna <- read_sf(dsn = here("data", "spatial", "PNA_EEZ"),
               layer = "PNA_EEZ") %>% 
  select(ISO_Ter3) %>% 
  st_simplify(11) %>% 
  st_rotate() %>% 
  st_union(by_feature = T) %>% 
  st_crop(y = c(xmin = 160, xmax = 210, ymin = -15, ymax = 10))

colors = c(
  "#0C276C",
  "#0B3B74",
  "#0A4E7C",
  "#096284",
  "#08768C",
  "#078994",
  "#059D9B",
  "#04B0A3",
  "#03C4AB",
  "#02D8B3",
  "#01EBBB",
  "#00FFC3"
)

ggplot() +
  geom_raster(data = rast,
              mapping = aes(x = x, y = y, fill = days)) +
  geom_sf(data = pna, fill = "transparent", color = "steelblue") +
  geom_sf(data = pipa, fill = "transparent", color = "red", size = 1) +
  facet_wrap(~post, ncol = 1) +
  # scale_fill_viridis_c() +
  scale_fill_gradientn(colors = colors) +
  guides(fill = guide_colorbar(title = quo(log[10]~"(days)"),
                               frame.colour = "white",
                               ticks.colour = "white")) +
  dark_theme_minimal() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(strip.text = element_text(size = 12)) +
  labs(x = "", y = "") +
  ggtitle(label = "Phoenix Islands Protected Area (PIPA), Kiribati")
```

## Scope and Motivation

- For the countries:

  - What are the costs associated to establishing an MPA?

  - What factors drive these costs?

  - How can these costs be reduced or eliminated?

- For the fishing vessels:

  - How do they respond?

  - Where do they go?

## Approach

- 10-patch model of the fishery (9 patches under VDS, one patch under Open Access)

- Vessel-level profits for fishing in patch $i$ are:

$$
\Pi_i(E_i,X_i,R) = pqE_iX_i\Omega_i-cE_i^\beta
$$

- In patch 1, $\Omega_1 = \theta + (1 - \theta)(1 - R)$

- $\Omega = 1$ for patches $i > 1$

- Vessel-day price must equate the marginal profits from the last unit of effort in patches $i = (1, 9)$

$$
\pi_i(E_i) = pqX_i\Omega_i- \beta cE_i^{\beta-1}
$$

## Approach

- Effort is capped at 45,000 vessel-days ($\bar{E}$)

- Measure country-level demand curves

  - No trading: price is patch-specific

  - Trading: price is the same for all patches

```{r}
knitr::include_graphics(path = here("docs", "slides", "img", "demand_curves.png"))
```

## Approach

## Less movement increases costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "cost_no_trading.png"))
```

## Vessel-day trading reduces costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "cost_with_trading.png"))
```

## Allocation rules matter

## What can we learn from PIPA

- Implemented in January, 2015

- 397,447 Km^2^

- 1.5 times area of Ecuador

- ~2% of total PNA area

- ~10% of fishing activity in PNA

## Generalizability to Palau