### 2019-05-04 scripts/model/2_run_model.Rmd
Moved modelling scripts to Rmd files

---
title: "Run the model"
author: "Villaseñor-Derbez J.C."
date: "Updated on `r Sys.Date()`"
output:
  html_document: 
    code_folding: show
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: false
---

# Setup

## Rmd options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE)

# Clean the environment
rm(list = ls())
```

## Load packages and functions

```{r}
# Load libraries
library(startR)
library(magrittr)
library(here)
library(cowplot)
library(furrr)
library(tidyverse)

# Load functions
source(here("scripts", "model", "grow.R"))
source(here("scripts", "model", "E_vec_fxn.R"))
source(here("scripts", "model", "pi_of_E.R"))
source(here("scripts", "model", "max_Ei.R"))
source(here("scripts", "model", "solve_pi.R"))
source(here("scripts", "model", "PNA_with_trading.R"))
source(here("scripts", "model", "PNA_no_trading.R"))

# Some housekeeping
update_geom_defaults("point",
                     list(
                       fill = "steelblue",
                       color = "black",
                       shape = 21,
                       size = 3
                     ))

update_geom_defaults("line", list(color = "black",
                                  size = 1))

theme_set(startR::ggtheme_plot())
```

## Define parameters

```{r}
# Reserve properties
theta <- 1
R <- 0

# Bio
MSY <- 1875600 # 50th percentile from MSY in table 8 of SA (https://www.wcpfc.int/node/27490)
Bmsy <- 1628000 # 50th percentile from SBmsy in table 8 of SA (https://www.wcpfc.int/node/27490)
K <- 6876526 # 50th percentile from SB_f0 in table 8 of SA (https://www.wcpfc.int/node/27490)
Bc_B0 <- 0.51 # 50th percentile from SBlatest/SB_fo in table 8 of SA (https://www.wcpfc.int/node/27490)
Cnow <- 1679444 # Catches now
Bnow <- K * Bc_B0 # current Biomass (2012 - 2015 average)
r <- 0.57 # From fishbase: Prior r = 0.57, 95% CL = 0.41 - 0.78 (https://www.fishbase.in/summary/107#)

# Economic
beta <- 1.3
p <- 1100 # (1447 + 1467) / 2 #mean between thailand and japan values (Value of WCPFC-CA tuna fisheries 2017 report)
E <- 45000
q <- 12 / (0.1 * Bnow) # 2 * (0.8 * Cnow) / (E * 0.8 * Bnow)
c <- 180 # 340

# Vectors of initial biomass and movement
# Proportional size of the high seas
prop_hs <- 0.1
# Number of patches
n_patches <- 10
# Biomass redistribution vector
fvec <- c(rep(((1 - prop_hs) / (n_patches - 1)), (n_patches - 1)), prop_hs)
# fvec <- c(0.12, 0.08, 0.07, 0.13, 0.07, 0.12, 0.09, 0.12, 0.1, 0.1)
names(fvec) <- c("KIR", "FSM", "MHL", "NRU", "PLW", "PNG", "SLB", "TKL", "TUV", "HS")
Xvec <- Bnow * fvec
```

# Cost of a reserve is governed by fish movement

> Less movement leads to higer costs

This first par uses no trading, but allows us to look at how movement ($\theta$) governs costs. The intuition is that, given the parameterization, high movement rates will lead to low costs because:

$$\Omega = \theta + (1 - \theta)(1 - R)$$

This shows that, no matter what the size of the reserve is, when $\theta = 1$, $\Omega = 1$.

## Get initial conditions

We first run a model that allows for trading to get the Effort vector that would maximize $pi$.

```{r}
Benchmark_results <-
  PNA_with_trading(
    fvec = fvec,
    theta = theta,
    R = R,
    r = r,
    K = K,
    p = p,
    q = q,
    Xvec = Xvec,
    beta = beta,
    c = c,
    E = E
  ) %>%
  select(
    Country,
    b_harvest = Harvest,
    b_vessel_days = Vessel_Days,
    b_stock = Stock,
    b_stock_i = Stock_i,
    b_vdsprice = VDSprice,
    b_vdsrevenue = VDSrevenue,
    b_vdsrevenue_all = VDSrevenue_all,
    b_vdsrevenue_notKIR = VDSrevenue_notKIR
  )
```

We then take these results, and use the observed efforts as the optimal efforts.

```{r}
Evec <- Benchmark_results$b_vessel_days

Evec
```

## Run the model

We now run the model with **NO** trading, using the Evec above as starting condition.

```{r}
# Create an empty tibble to store the results in the process
Rvec <- seq(0, 1, 0.05)
thetavec <- seq(0, 1, 0.1)

DF_results_no_trading <-
  expand.grid(R = Rvec, theta = thetavec) %>%
  mutate(
    res = map2(
      theta,
      R,
      PNA_no_trading,
      fvec = fvec,
      r = r,
      K = K,
      p = p,
      q = q,
      Xvec = Xvec,
      beta = beta,
      c = c,
      Evec = Evec
    )
  ) %>%
  unnest()
```

We now take these results and add the benchmark results (no reserve) to calculate the cost of the reserve as the change in revenue. The proportional way of saying this is:

$$
\%C_R = \frac{\Pi_{R = 0} - \Pi_{R = r}}{\Pi_{R = 0}}
$$
Where $\Pi_{R = 0}$  is the vessel-day revenue when there is no reserve, hence the $R = 0$ subindex. The $\Pi_{R = r}$ denotes the vessel-day revenues when reserve has size ($r \in (0, 1)$). The vessel-day revenue is simply the vessel-day price times the number of vessel-days in patch $i$: $\pi \times E_i$.

```{r}
DF_1_no_trading <- DF_results_no_trading %>%
  filter(Country == "KIR") %>%
  left_join(Benchmark_results, by = "Country") %>%
  mutate(
    cost = b_vdsrevenue - VDSrevenue,
    rel_cost = cost / b_vdsrevenue
  )
```

We can now go ahead and plot these results. The figure below shows the $C_R$ defined above for different size of reserves $R$ in the x-axis, given values of $theta$ (colors). The graph shows how a smaller movement rate leads to increased costs to the country that implements the reserve. A movement of $\theta = 1$ implies no reserve, and thus leads to no losses. However, this parameter is unlikely. We also see how the losses are not proportional to 1. That is, a 10% closure does not lead to a 10% loss in profits.

**Is this due to the concavity in costs? i.e. the $\beta$ parameter$?**

Yes, it looks like a $\beta = 1.1$ actually reduces this.

```{r, fig.width = 4, fig.height = 4}
(
  costs_no_trading_plot <-
    ggplot(data = DF_1_no_trading) +
    geom_line(aes(
      x = R,
      y = rel_cost,
      group = theta,
      color = theta
    ),
    size = 1
    ) +
    xlab("Portion of patch as reserve (R)") +
    ylab("Costs (% of case with no reserve)") +
    scale_color_viridis_c() +
    theme(
      text = element_text(size = 10),
      axis.text = element_text(size = 8)
    ) +
    startR::ggtheme_plot() +
    scale_y_continuous(labels = scales::percent) +
    guides(color = guide_colorbar(title = quo("Movement ("~theta~")"),
                                  ticks.colour = "black", frame.colour = "black")) +
    theme(legend.justification = c(0, 0.9),
          legend.position = c(0, 1)) +
    geom_abline(intercept = 0, slope = 1, size = 1, linetype = "dashed", color = "red")
)
```

```{r}
ggsave(plot = costs_no_trading_plot,
       filename = here("docs", "slides", "img", "cost_no_trading.png"),
       width = 3,
       height = 3)
```

# Trading affects costs

> Trading reduces costs to implementing party

We now run the same model, with same initial conditions, but allow trading as a reserve is closed.

```{r}
DF_results_with_trading <-
  expand.grid(R = Rvec, theta = thetavec) %>%
  mutate(
    res = map2(
      theta,
      R,
      PNA_with_trading,
      fvec = fvec,
      r = r,
      K = K,
      p = p,
      q = q,
      Xvec = Xvec,
      beta = beta,
      c = c,
      E = E
    )
  ) %>%
  unnest()
```

Again, put it together with the reference values to calculate costs.

```{r}
DF_1_with_trading <- DF_results_with_trading %>%
  filter(Country == "KIR") %>%
  left_join(Benchmark_results, by = "Country") %>%
  mutate(
    vessel_days_sold = b_vessel_days - Vessel_Days,
    sales_revenue = vessel_days_sold * VDSprice,
    all_VDSrevenue = VDSrevenue + sales_revenue,
    cost = b_vdsrevenue - all_VDSrevenue,
    rel_cost = cost / b_vdsrevenue
  )
```

The following graph shows the same information as the graph above. However, in this case, the costs are significantly lower. This is because a party can sell it's original allocation to other parties. The reduction i

```{r, fig.width = 4, fig.height = 4}
(costs_with_trading_plot_long_axis <- ggplot(data = DF_1_with_trading) +
  geom_line(aes(
    x = R,
    y = rel_cost,
    group = theta,
    color = theta
  ),
  size = 1
  ) +
  xlab("Portion of patch as reserve (R)") +
  ylab("Costs (% of case with no reserve)") +
  scale_color_viridis_c() +
  startR::ggtheme_plot() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme(legend.position = "None"))
```

```{r}
ggsave(plot = costs_with_trading_plot_long_axis,
       filename = here("docs", "slides", "img", "cost_with_trading_long_axis.png"),
       width = 3,
       height = 3)
```

```{r}
(costs_with_trading_plot <- ggplot(data = DF_1_with_trading) +
  geom_line(aes(
    x = R,
    y = rel_cost,
    group = theta,
    color = theta
  ),
  size = 1
  ) +
  xlab("Portion of patch as reserve (R)") +
  ylab("Costs (% of case with no reserve)") +
  scale_color_viridis_c() +
  startR::ggtheme_plot() +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "None"))
```

```{r}
ggsave(plot = costs_with_trading_plot,
       filename = here("docs", "slides", "img", "cost_with_trading.png"),
       width = 3,
       height = 3)
```

## Cost reduction due to trading

```{r}
shorter_no_trading <- DF_1_no_trading %>% 
  select(Country, R, theta, Vessel_Days, VDSprice, VDSrevenue, cost, rel_cost)

shorter_trading <- DF_1_with_trading %>% 
  select(Country, R, theta, Vessel_Days_trading = Vessel_Days, pi_trading = VDSprice, VDSrevenue_trading = VDSrevenue, cost_trading = cost, rel_cost_trading = rel_cost)

both <- left_join(shorter_no_trading, shorter_trading, by = c("Country", "R", "theta")) %>% 
  mutate(delta_cost = cost - cost_trading,
         rel_delta_cost = delta_cost / cost) %>% 
  filter(!R == 0,
         !theta == 1)
```

```{r}
(delta_costs_plot <- ggplot(data = both) +
  geom_line(aes(
    x = R,
    y = rel_delta_cost,
    group = theta,
    color = theta
  ),
  size = 1
  ) +
  xlab("Portion of patch as reserve (R)") +
  ylab("Costs avoided by trading") +
  scale_color_viridis_c(limits = c(0, 1)) +
  theme(
    text = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) +
  startR::ggtheme_plot() +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "None"))
```

All together

```{r, fig.width = 4, fig.height = 7}
(cost_plot <- cowplot::plot_grid(costs_no_trading_plot,
                   costs_with_trading_plot,
                   delta_costs_plot, ncol = 1))
```

Now let's save this figure.

```{r}
ggsave(plot = cost_plot,
       filename = here("docs", "img", "cost_plot.pdf"), width = 4, height = 7)
```

## How does vessel-price change?

### When there's no trading

```{r}
DF_results_no_trading %>% 
  filter(!Country == "HS") %>% 
  mutate(Country = fct_relevel(Country, "KIR")) %>% 
  ggplot() +
  geom_line(aes(
    x = R,
    y = VDSprice,
    group = theta,
    color = theta
  ),
  size = 1
  ) +
  xlab("Portion of patch as reserve (R)") +
  ylab(quo("Vessel-day Price ("~pi~")")) +
  scale_color_viridis_c() +
  theme(
    text = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) +
  startR::ggtheme_plot() +
  guides(color = guide_colorbar(title = quo("Movement ("~theta~")"),
                                ticks.colour = "black",
                                frame.colour = "black")) +
  facet_wrap(~Country, scales = "free_y")
```

### When there is trading

```{r}
ggplot(data = DF_1_with_trading) +
  geom_line(aes(
    x = R,
    y = VDSprice,
    group = theta,
    color = theta
  ),
  size = 1
  ) +
  xlab("Portion of patch as reserve (R)") +
  ylab(quo("Vessel-day Price ("~pi~")")) +
  scale_color_viridis_c() +
  theme(
    text = element_text(size = 10),
    axis.text = element_text(size = 8)
  ) +
  startR::ggtheme_plot() +
  guides(color = guide_colorbar(title = quo("Movement ("~theta~")"),
                                ticks.colour = "black",
                                frame.colour = "black"))
```

# Allocation rules matter

```{r}
source(here("scripts", "model", "allocate.R"))
source(here("scripts", "model", "allocate_wrap.R"))
source(here("scripts", "model", "move_in_time.R"))
source(here("scripts", "model", "move_in_time_no_trading.R"))
```

We first obtain the optimal allocation by running the model until steady state and with no reserves.

```{r}

alphas <- c(0, 0.25, 0.5, 0.75, 1)

delta <- 1 / 1.1

plan(multiprocess)

allocation <-
  expand.grid(
    R = Rvec,
    theta = c(0, 0.5, 1),
    alpha = alphas,
    trade = c(T, F)
  ) %>%
  as_tibble() %>%
  mutate(
    resutls = future_pmap(
      .l = list(
        alpha = alpha,
        R = R,
        theta = theta,
        trade = trade
      ),
      .f = allocate_wrap, 
      fvec = fvec,
      r = r,
      K = K,
      p = p,
      q = q,
      Xvec = Xvec,
      beta = beta,
      c = c,
      E = E
    )
  ) %>%
  unnest() %>%
  filter(country == "KIR") %>%
  mutate(
    rev_dir = Ei * pi,
    deltaE = Ea - Ei,
    rev_sells = deltaE * pi,
    rev_tot = rev_dir + rev_sells,
    fct = delta ^ year,
    rev_tot_disc = rev_tot * fct
  )

best_rev <- allocate_wrap(
  alpha = 0,
  R = 0,
  theta = 1,
  trade = T,
  fvec = fvec,
  r = r,
  K = K,
  p = p,
  q = q,
  Xvec = Xvec,
  beta = beta,
  c = c,
  E = E
) %>%
  filter(country == "KIR") %>%
  mutate(
    rev_dir = Ei * pi,
    deltaE = Ea - Ei,
    rev_sells = deltaE * pi,
    rev_tot = rev_dir + rev_sells,
    fct = delta ^ year,
    rev_tot_disc = rev_tot * fct
  ) %$%
  sum(rev_tot_disc)
```

```{r}
(allocation_cost_plot <- allocation %>%
  mutate(trade = ifelse(trade, "Trading", "No Trading"),
         theta = as.character(theta)) %>% 
  group_by(trade, alpha, R, theta) %>%
  summarize(rev = sum(rev_tot_disc)) %>%
  ungroup() %>%
  mutate(cost = (best_rev - rev) / best_rev) %>%
  ggplot(aes(x = R, y = cost, color = alpha, group = paste(alpha, theta), linetype = theta)) +
  geom_line(size = 1) +
  scale_y_continuous(labels = scales::percent) +
  xlab("Portion of patch as reserve (R)") +
  ylab("Costs (% of case with no reserve)") +
  scale_color_viridis_c() +
  guides(color = guide_colorbar(ticks.colour = "black", frame.colour = "black")) +
  ggtheme_plot() +
  facet_wrap(~trade) +
  guides(color = guide_colorbar(title = quo("Allocation\nparameter ("~alpha~")")),
         linetype = guide_legend(title = quo("Movement\nparameter ("~theta~")"))))
```

```{r}
ggsave(
  plot = allocation_cost_plot,
  filename = here("docs", "slides", "img", "allocation_cost_plot.pdf"),
  width = 5,
  height = 3
)
```

```{r}
allocation %>%
  filter(!trade) %>% 
  ggplot(aes(x = year, y = Ea, color = R, group = paste(alpha, R, theta))) +
  geom_line(size = 1) +
  facet_grid(alpha ~ theta) +
  scale_color_viridis_c() +
  guides(color = guide_colorbar(ticks.colour = "black", frame.colour = "black")) +
  ggtheme_plot()
```

```{r}
allocation %>%
  filter(theta == 0.5, !trade) %>% 
  ggplot(aes(x = year, y = Xi, color = R, group = paste(alpha, R))) +
  geom_line(size = 1) +
  facet_grid(~alpha) +
  scale_color_viridis_c() +
  guides(color = guide_colorbar(ticks.colour = "black", frame.colour = "black")) +
  ggtheme_plot()
```

# Reproducibility chunk

```{r}
sessionInfo()
```