### 2019-05-14 docs/slides/NCEAS_presentation.Rmd
Added and updated slides for NCEAS presentation last week

---
title: "Marine Conservation and Fishing Effort Markets[^*]"
author:
  - Villaseñor-Derbez J.C.$^1$
  - Lynham J.$^2$
  - Costello C.$^{1, 3}$
institute: 
   - $^1$Bren School, UCSB
   - $^2$Department of Economics, UHM
   - $^3$emLAB, UCSB
subtitle: "NCEAS-Future Earth Working Group"
date: "May 6, 2019"
output: 
  beamer_presentation: 
    theme: Darmstadt
header-includes:
  - \usepackage{media9}
  - \usepackage{eso-pic,graphicx,transparent}
---

[^*]: This is work in progress developed for the NCEAS working group. Please do not share externally. Results are yet to be refined and peer-reviewed.

\AddToShipoutPictureFG{
  \AtPageCenter{% or \AtTextCenter
    \makebox[0pt]{\rotatebox[origin=c]{45}{%
      \scalebox{5}{\texttransparent{0.3}{DRAFT}}%
    }}
  }
}

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE#,
                      # fig.align = "center"
                      )

suppressPackageStartupMessages({
  library(here)
})
```

## Outline

- Background and motivation

- Scope of the project

- Approach

  - Model and assumptions

  - Phoenix Island Protected Area*

- Generalizability to Palau

# Background and motivation

## Motivation

\begin{center}
	\includemedia[
    label=video1,
    width=300pt,height=225pt,%width=300pt,height=225pt,
	]{}{GFW_video.swf}

    \mediabutton[mediacommand=video1:playPause,
    overface=\color{blue}{\fbox{\strut Play/Pause}},
    downface=\color{red}{\fbox{\strut Play/Pause}}
    ]{\fbox{\strut Play/Pause}}
\end{center}

## Motivation

### Effort displacement by large-scale MPAs

- Where was effort displaced to?

- Does effort displacement imply revenue losses from VDS?

- Does a 30\% closure mean a 30\% loss in revenue?

# Scope and Approach

## Scope

Can economic instruments incentivize conservation?

- For the countries:

  - What are the costs associated to establishing an MPA?

  - What factors drive these costs?

  - Can these costs be reduced or eliminated?

- For the fishing vessels:

  - How do they respond?

  - Where do they go?

## Approach

- 10-patch model of the tuna purse-seine fishery

  - 9 patches under VDS
  - one patch under Open Access

- Profits from fishing in patch $i$ are:

$$\Pi_i(E_i,X_i,R) = pqE_iX_i\Omega_i-cE_i^\beta$$

$$\Omega_1 = \theta + (1 - \theta)(1 - R)$$

- Vessel-day price must equate the marginal profits from the last unit of effort in patches $i = (1, 9)$

$$\pi_i(E_i) = pqX_i\Omega_i- \beta cE_i^{\beta-1}$$

## Approach

- Patch-level effort is then

$$E_i = \left(\frac{pqX_i\Omega_i - \pi_i}{\beta c }\right) ^ { \frac{1}{\beta - 1}}$$

- Effort is capped at 45,000 vessel-days ($\bar{E}$)

$$\bar{E} = \sum_{i = 1}^9\left(\frac{pqX_i\Omega_i - \pi}{\beta c }\right) ^ {\frac{1}{\beta - 1}}$$

## Approach

- Measure country-level demand curves

  - No trading: price is patch-specific

  - Trading: price is the same for all patches

```{r}
# Recortar para que salgan hasta 3000 y 10000 solamente
knitr::include_graphics(path = here("docs", "slides", "img", "demand_curves.png"))
```

# Model predictions

## Model predictions

### Relationship between closure size and losses

Does a 30\% closure mean a 30\% loss in revenue?

## Losses with no trading

```{r}
knitr::include_graphics(here("docs", "slides", "img", "cost_no_trading.png"))
```

## Model predictions

### A market for fishing efforts

- What is the effect of trading?

- Can costs be reduced?

## Trading significantly reduces costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "cost_with_trading_long_axis.png"))
```

## Trading significantly reduces costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "cost_with_trading.png"))
```

## Allocation rules matter

- Vessel-days are allocated to countries each year

- Combination of biomass and in-EEZ usage

$$
E_i^* = \alpha \left(\frac{\sum_{\tau = 0}^{\hat{\tau}}E_{i,t-\tau}}{\bar{E}\hat{\tau}} \right) +
(1 - \alpha) \left(\frac{\sum_{\tau = 0}^{\hat{\tau}}X_{i,t-\tau}}{\bar{X}\hat{\tau}} \right)
$$

- $\alpha$ is a weight on historical effort ($E_i$) and historical biomass ($B_i$)

What is the importance of allocation rules?

## "*Biomass-based*" allocation reduces costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "allocation_cost_plot.pdf"))
```

# Generalizability to Palau

## Palau Vessel-day schemes

- Purse seine vessel-days

  - 700 purse seine vessel-days

  - \$5.60 - \$8.75 Million USD

  - Tradable

  - Trading might reduce costs, but allocation rules matter

- Longline vessel-days

  - 10,500 longline vessel-days

  - \$2.10 Million USD

  - Non-tradable

## Conclusions

- Under no trading, implementing a reserve is costly

  - More fish movement reduces costs

- Trading lessens the blow of a spatial closure

  - \% costs avoided by trading decrease as biomass in implementing patch increases

- Allocation rules matter

  - A biomass-based allocation ($\alpha = 0$) reduces long-term costs

# Extra slides

## GFW summary for Palau

```{r}
knitr::include_graphics(here("docs", "slides", "img", "vessels_in_plw.png"))
```

## What can we learn from PIPA?

PIPA:

- Implemented in January, 2015

- 397,447 $Km^2$

- 2.7\% of total PNA area

- 11\% of Kiribati EEZ

- 22\% of Kiribati EEZ (excluding Line Islands)

GFW data:

- 313 tuna purse seine vessels that fished in PNA

- 92 Fished both before and after PIPA implementation:

  - 64 Fished within PIPA before implementation

- 2012 - Present

## Two "fleets"

```{r}
knitr::include_graphics(here("docs", "slides", "img", "sample_tracks.png"))
```

## Effort displacement in Kiribati

```{r}
knitr::include_graphics(here("docs", "img", "all_PS_VDS_KIR_year.pdf"))
```

## Effort displacement in all PNA

```{r}
knitr::include_graphics(here("docs", "img", "all_PS_VDS_year.pdf"))
```

## Displaced vessels

```{r}
knitr::include_graphics(here("docs", "img", "fishing_raster_displaced.png"))
```

## Non-displaced vessels

```{r}
knitr::include_graphics(here("docs", "img", "fishing_raster_not_displaced.png"))
```

## Difference

```{r}
knitr::include_graphics(here("docs", "img", "fishing_raster_difference_between_groups.png"))
```

# GFW Data

## Effort vs catches

```{r}
knitr::include_graphics(here("docs", "slides", "img", "catch_effort_plot.png"))
```

## Effort and revenues

```{r}
knitr::include_graphics(here("docs", "img", "revenue_FFA_GFW_linear.pdf"))
```