### 2019-05-14 docs/tab/parameters.tex
Added new MS

\begin{tabular}{l|r|l}
\hline
Parameter & Value & Source\\
\hline
MSY & 1.875600e+06 & 50th percentile from MSY in table 8 of Stock Assessment\\
\hline
$B_{msy}$ & 1.628000e+06 & 50th percentile from MSY in table 8 of Stock Assessment\\
\hline
K & 6.876526e+06 & 50th percentile from MSY in table 8 of Stock Assessment\\
\hline
$B_c/B_{msy}$ & 5.100000e-01 & 50th percentile from MSY in table 8 of Stock Assessment\\
\hline
$C_{now}$ & 1.679444e+06 & Catches from Stock Assessment\\
\hline
$B_{now}$ & 3.507028e+06 & Current Biomass (2012 - 2015 average)\\
\hline
$r$ & 5.700000e-01 & From fishbase: Prior r = 0.57, 95 CL = 0.41 - 0.78\\
\hline
$\beta$ & 1.300000e+00 & Standard\\
\hline
p & 1.100000e+03 & Mean between Thailand and Japan values (Value of WCPFC-CA tuna fisheries 2017 report)\\
\hline
q & 3.420000e-05 & Estimated so that efforts match catches given biomass and vessel-day prices\\
\hline
c & 1.800000e+02 & -\\
\hline
f & 1.000000e-01 & Biomass is equally distributed between patches ($f_i = 0.1$)\\
\hline
\end{tabular}