### 2019-10-15 docs/tab/balance_table.tex
Updated tables

\begin{table}[t]

\caption{\label{tab:}Mean values on observable characteristics by vessel for displaced (n = 64), and non-displaced vessels (n = 254). Numbers in parentheses indicate standard deviation. The last column contains the difference in means (t-scores), with asterisks inicating significant differences as indicated by a two-tailed t-test (* p < 0.1; ** p< 0.05; *** p < 0.01).}
\centering
\begin{tabular}{l|l|l|l}
\hline
Characteristic & Non-displaced & Displaced & Difference\\
\hline
Crew size (n) & 30.46 (6.25) & 26.38 (3.94) & -4.08 (-6.49) ***\\
\hline
Engine Power (KW) & 2559.89 (588.28) & 2983.6 (558.76) & 423.71 (5.36) ***\\
\hline
Length (m) & 68.97 (8.42) & 74.23 (9.71) & 5.25 (3.97) ***\\
\hline
Tonnage (GT) & 1383.41 (533.56) & 1718.14 (653.38) & 334.73 (3.79) ***\\
\hline
\end{tabular}
\end{table}